\section{CPU Scheduling and Synchronization}\label{sec:CPU_Scheduling_Synchronization}
The growing importance of multicore systems has brought an increased emphasis on developing multithreaded applications.
In such applications, several threads, which may be sharing data, are running in parallel on different processing cores.
These \nameref{def:Process}es are called \nameref{def:Cooperating_Process}es.

\begin{definition}[Cooperating Process]\label{def:Cooperating_Process}
  A \emph{cooperating process} is one that can affect or be affected by other \nameref{def:Process}es executing on a system.
  They can share a logical address space (code and data), \textbf{\nameref{def:Thread}s}, or can share data through files and/or messages, \textbf{\nameref{subsubsec:Communications}}.
\end{definition}

Given the way that multiple \nameref{def:Thread}s can be scheduled, namely in any order (relatively speaking), as programmers, we cannot be certain about which thread will be scheduled first.
This leads to all sorts of problems because of sharing information between multiple users.
The largest, and likely the most common, error in a multi\nameref{def:Thread}ed program is the \nameref{def:Race_Condition}.

\begin{definition}[Race Condition]\label{def:Race_Condition}
  A \emph{race condition} is when several processes access and manipulate the same data concurrently and the outcome of the execution depends on the particular order in which the access takes place.
  The only way to prevent a race condition is to ensure that \textbf{only one \nameref{def:Thread} can change the value at a time}.
\end{definition}

\input{./EDAF35-Operating_Systems-Sections/CPU_Scheduling_Synchronization/Synchronization}

\input{./EDAF35-Operating_Systems-Sections/CPU_Scheduling_Synchronization/Scheduling}

\input{./EDAF35-Operating_Systems-Sections/CPU_Scheduling_Synchronization/Thread_Scheduling}

\begin{definition}[Deadlock]\label{def:Deadlock}
  \emph{Deadlock} is when 2 processes require information from each other to continue running.
  If this happens, neither process will provide the other with its required information, so they will both wait for each other, forever.
\end{definition}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
