\section{File-System Interface}\label{sec:FS_Interface}
So that the computer system will be convenient to use, the \nameref{def:Operating_System} provides a uniform logical view of stored information, the \nameref{def:File}.

\defFile{}

Files in all \nameref{def:Operating_System}s have attributes that help describe them.
\begin{itemize}[noitemsep]
\item \textbf{Name}.
  The symbolic file name is the only information kept in human-readable form.
\item \textbf{Identifier}.
  This unique tag, usually a number, identifies the file within the file system; it is the non-human-readable name for the file.
\item \textbf{Type}.
  This information is needed for systems that support different types of files.
\item \textbf{Location}.
  This information is a pointer to a device and to the location of the file on that device.
\item \textbf{Size}.
  The current size of the file and possibly the maximum allowed size are included in this attribute.
\item \textbf{Protection}.
  Access-control information determines who can do reading, writing, executing, etc.
\item \textbf{Time, date, and user identification}.
  This information may be kept for creation, last modification, and last use.
  This data can be useful for protection, security, and usage monitoring.
\end{itemize}

The information about all files is kept in the directory structure, which is also on secondary storage.
An entry in a directory consists of the file’s name and its unique identifier.
The file's identifier in turn locates the other file attributes.
It may take more than a kilobyte to record this information for each file.
Because directories must also be nonvolatile, they must be stored on the device and brought into memory as needed.

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/File_Operations}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/File_Types}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/File_Structure}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/Access_Methods}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/Directory_Disk_Structure}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/FS_Mounting}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/File_Sharing}

\input{./EDAF35-Operating_Systems-Sections/FS_Interface/Protection}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
