\section{Main Memory}\label{sec:Main_Memory}
To truly benefit from \nameref{sec:CPU_Scheduling_Synchronization}, discussed in \Cref{sec:CPU_Scheduling_Synchronization}, we must be able to keep multiple \nameref{def:Process}es in memory at the same time.

In a \nameref{def:CPU}, the only things that can be accessed directly are \nameref{def:Register}s and main memory.
The CPU can reach the registers within one clock cycle, but accessing main memory takes several clock cycles, because access is done through the memory bus.
To prevent \nameref{def:Memory_Stall}s, we can use multiple \nameref{def:Thread}s, and we can add additional \nameref{def:Cache}s to the CPU itself.
This \nameref{def:Hardware} automatically speeds up memory access, without \nameref{def:Operating_System} intervention.

\begin{definition}[Cache]\label{def:Cache}
  A \emph{cache} is a very small amount of memory, built onto the \nameref{def:CPU} itself, and acts as a buffer between the CPU and main memory.
  A cache will have a copy of a small portion of what is in main memory, and the CPU goes to find the next thing, whatever it may be, from the cache first.
  Because the cache sits on the CPU itself, and does not have to interact with the memory bus, it is significantly faster than accessing main memory, but still slightly slower than accessing a register.
\end{definition}

Not only do we want our memory and its access to be fast, we also want to make sure that the memory and its contents are used correctly.
For proper system operation we must protect the operating system from access by user processes.
On multiuser systems, we must additionally protect user processes from one another.
This protection must be provided by the hardware because the operating system doesn’t usually intervene between the CPU and its memory accesses (because of the resulting performance penalty).
A \nameref{def:Memory_Management_Unit} implements this production, in hardware, in several different possible ways.

\begin{definition}[Memory Management Unit]\label{def:Memory_Management_Unit}
  The \emph{Memory Management Unit} (\emph{MMU}) is a piece of hardware built onto a \nameref{def:CPU} (See \Cref{rmk:MMU_Location}).
  It is the main controller for all main memory accesses made by a CPU.\@

  It handles the run-time mapping from \nameref{def:Virtual_Address}es to \nameref{def:Physical_Address}es.

  \begin{remark}[MMU Location]\label{rmk:MMU_Location}
    The location of the \nameref{def:Memory_Management_Unit} is \textbf{typically} on the CPU.\@
    This holds true for modern CPUs, however, older CPUs like the Motorola 68000 series have physically separate MMUs, which can be attached to the motherboard that the CPU sits on.
  \end{remark}
\end{definition}

To ensure that everything works, the first thing that needs to be done is to separate each \nameref{def:Process} into its own memory space.
This protects the processes from each other and is fundamental to having multiple processes loaded in memory for concurrent execution.
To separate memory spaces, we need the ability to determine the range of legal addresses that the process may access and to ensure that the process can access only these legal addresses.

\input{./EDAF35-Operating_Systems-Sections/Main_Memory/Address_Binding}

\input{./EDAF35-Operating_Systems-Sections/Main_Memory/Logical_Physical_Virtual_Address_Spaces}

\input{./EDAF35-Operating_Systems-Sections/Main_Memory/Dynamic_Loading}

\input{./EDAF35-Operating_Systems-Sections/Main_Memory/Dynamic_Linking}

\input{./EDAF35-Operating_Systems-Sections/Main_Memory/Swapping}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
