\section{Input/Output (I/O)}\label{sec:IO}
In \textsc{unix}, I/O devices include:
\begin{itemize}[noitemsep]
\item Disk
\item Terminal
\item Shared Memory
\item Printer
\item Network
\end{itemize}

This is mostly because \textsc{unix} made the design decision to try to view every component of a system as a \nameref{def:File}.

Due to the variety of I/O devices that need to be supported, there are a vast number of different mechanisms for using these devices.
But, there are a few common mechanisms, requirements, and activities:
\begin{itemize}[noitemsep]
\item Read/Write Ops
\item Metadata:
  \begin{itemize}[noitemsep]
  \item Name
  \item Position
  \item Directory Name
  \item Creation Date
  \item Last Access Date
  \item IP Address
  \item MAC Address
  \item TCP Packet Sequence Number
  \end{itemize}
\item Robustness
\item Thread-safety
\end{itemize}

There are few general concerns that we need to have about the idea of viewing everything as a \nameref{def:File}.
\begin{itemize}[noitemsep]
\item How are I/O endpoints represented?
  \begin{itemize}[noitemsep]
  \item \nameref{def:File_Descriptor}
  \end{itemize}
\item How do we perform I/O?\@
  \begin{itemize}[noitemsep]
  \item Byte at a time
  \item Give a chunk of memory and later check that the requested I/O completed?
  \end{itemize}
\item How do we perform I/O \emph{efficiently}?
  \begin{itemize}[noitemsep]
  \item Efficiency depends on what we define efficient to be. Essentially, what are we optimizing for?
  \end{itemize}
\end{itemize}

\subsection{I/O Devices}\label{subsec:IO_Devices}
There are 2 major types of I/O devices on \textsc{unix} systems:
\begin{enumerate}[noitemsep]
\item \nameref{def:Block_Device}s
\item \nameref{def:Character_Device}s
\end{enumerate}

\begin{definition}[Block Device]\label{def:Block_Device}
  A \emph{block device} is an I/O device that accesses and stores data in fixed-sized blocks.
  Typically, this also means they have fixed total size as well.
  This means they support seeking through their contents and random access for parts of their contents.

  Some typical devices classified as this are:
  \begin{itemize}[noitemsep]
  \item Disk
  \item Memory
  \end{itemize}
\end{definition}

\begin{definition}[Character Device]\label{def:Character_Device}
  A \emph{character device} is an I/O device that access and receives data as a stream.
  This means it receives ``characters'' as a stream, one-by-one.
  There is no support for seeking or random access of the stream, because we are getting the data as soon as it is being given.

  Some typical devices in this category are:
  \begin{itemize}[noitemsep]
  \item Network
  \item Mouse
  \item Keyboard
  \end{itemize}
\end{definition}

\subsection{Filesystem}\label{subsec:Filesystem}
The filesystem acts as a namespace for devices, and allows for the efficient storage of data on \nameref{def:Block_Device}s.
A typical file system consists of two types of files:
\begin{enumerate}[noitemsep]
\item \textit{Regular files} consist of ASCII or binary data
  \begin{itemize}[noitemsep]
  \item Directories
  \end{itemize}
\item \textit{Special Files} may represent:
  \begin{itemize}[noitemsep]
  \item In-memory structures
  \item Sockets
  \item Raw Devices
  \end{itemize}
\end{enumerate}

\subsection{Files}\label{subsec:Files}
\begin{definition}[File]\label{def:File}
  A \emph{file} is an operating system abstraction over some other data.
  It allows us to interact with many different file systems and devices over a variety of protocols in a abstract and concise way.
  A file can be accessed by using a \textit{fully qualified path}.

  The only thing each file \textbf{MUST} have is a unique \nameref{def:inode}.
\end{definition}

\begin{definition}[\texttt{inode}]\label{def:inode}
  The \emph{inode} is a filesystem-unique number (Typically, there is one filesystem per device, so this is typically a per-device-unique number).
  The inode tracks:
  \begin{itemize}[noitemsep]
  \item Ownership
  \item Permissions
  \item Size
  \item Type
  \item Location
  \item Number of \nameref{def:Hard_Link}s
  \end{itemize}
\end{definition}

\begin{definition}[Hard Link]\label{def:Hard_Link}
  A \emph{hard link} is a link between \nameref{def:inode}s.
  Thus, they each point to the same data, and must have the same name.
  When one of the links is deleted, the total count for that inode decreases.
  Once the inode reaches \texttt{0} hard links, it is removed (deleted) from the system.
\end{definition}

\begin{definition}[Symlink]\label{def:Symlink}\label{def:Symbolic_Link}\label{def:Soft_Link}
  A \emph{symlink} (\emph{symbolic link}, \emph{soft link}) is a link between \nameref{def:File}s.
  Thus, the link points to the file, which then points to the data.
  The link is not required to have the same name as the original file.
  However, if the file that the symlink is pointing to is deleted, then we are left with a dangling pointer.
\end{definition}

However, the items discussed above are held strictly in storage, hard drive, meaning they aren't good for regular use, as they are too slow.
\begin{remark*}
  Barring you doing something like setting up a portion of your RAM as a ``hard drive'', allowing for ridiculous performance.
\end{remark*}

Thus, we load and open an \nameref{def:inode} into \nameref{def:Memory}, and call this copy a \nameref{def:vnode}.

\begin{definition}[\texttt{vnode}]\label{def:vnode}
  A \emph{vnode} is a copy of the \nameref{def:inode} of the current file.
  Every currently open file has a \textbf{single} \texttt{vnode}.
\end{definition}

Now that we have a way to efficiently access a file in-\nameref{def:Memory}, using the \nameref{def:vnode} structure, how can we ``open'' the same \nameref{def:File} multiple times?
This is handled by the \nameref{def:File_Description} structure.

\begin{definition}[File Description]\label{def:File_Description}
  The \emph{file description} allows the kernel to track which \nameref{def:Process} has opened which \nameref{def:File} and trace them back to their \nameref{def:vnode}.
  For \textbf{each process}, the kernel maintains a table of pointers to its open file structures.
  This table points to each file descriptor, which then point to the backing \nameref{def:vnode}s.

  The file description tracks:
  \begin{itemize}[noitemsep]
  \item Position
  \item Access Mode
  \item Pointer to the backing vnode
  \end{itemize}

  There is an open file description for \textbf{each occurrence} of a \nameref{def:File}'s opening.
\end{definition}

However, all of the structures discussed above are in \textbf{kernel memory}, meaning user \nameref{def:Process}es cannot interact with them.
This is where the \nameref{def:File_Descriptor} comes in.

\begin{definition}[File Descriptor]\label{def:File_Descriptor}
  The \emph{file descriptor} is the index of this file in the \nameref{def:File_Description} table.
  There are always three file descriptors defined at the beginning of a \nameref{def:Program}'s execution.
  \begin{enumerate}[noitemsep]
  \item FD \texttt{0} is STanDard INput (\texttt{STDIN})
  \item FD \texttt{1} is STanDard OUTput (\texttt{STDOUT})
  \item FD \texttt{2} is StanDard ERRor (\texttt{STDERR})
  \end{enumerate}
\end{definition}

After opening a file, \textbf{all} file operations are performed using \nameref{def:File_Descriptor}s.
This obscures kernel I/O and filesystem implementation details from the user, allowing for an elegant and abstract I/O API.\@

\subsection{System-Level I/O API}\label{subsec:System_Level_IO_API}
Input and output is one of the basic operations that a program of any use will have to do.
Below, there is a list of common \nameref{def:Syscall}s for performing these operations.
These \nameref{def:Syscall}s are the lowest-level I/O calls we can make for files

\begin{description}[noitemsep]
\item[\cinline{int open(const char *path, int oflag, ...)}] Opens a \nameref{def:File}.
  \begin{itemize}
  \item Loads \nameref{def:vnode} for \nameref{def:File} at \texttt{path}
  \item \texttt{oflag} is a bitwise OR of \cinline{O_RDONLY}, \cinline{O_WONLY}, \ldots{}
  \item Creates and initializes a new \nameref{def:File_Description} in the table
  \item Returns the first unused \nameref{def:File_Descriptor} available
  \item If you open the same file twice, then you get a new file descriptor (a new file description is made), but it points to the same \nameref{def:vnode}
  \item Process inherits parent's open files across a \cinline{fork}
  \item Process retains them after an \cinline{exec}
  \item Parent and child share:
    \begin{itemize}
    \item File position
    \item File Access mode
    \end{itemize}
  \item Sharing this file description allows for coordinating between separate process
  \item You can mirror this inside of a single process by using the \texttt{dup} \nameref{def:Syscall}.
  \end{itemize}
\item[\cinline{int fstat(int fd, struct stat *buf)}] Query for file metadata
  \begin{itemize}
  \item \cinline{struct stat fstat.st_ino} Get inode number
  \item \cinline{struct stat fstat.st_size} Get file size
  \item \cinline{struct stat fstat.st_nlink} Get number of hard links
  \end{itemize}
\item[\cinline{int dup(int fd)}] Duplicate the given \nameref{def:File_Descriptor} and return a new file descriptor that points to the same \nameref{def:File}.
\item[\cinline{int dup2(int fd1, int fd2)}] Duplicate \texttt{fd1} such that \texttt{fd2} writes to the same \nameref{def:vnode}.
  \begin{itemize}[noitemsep]
  \item We can use this to change the \nameref{def:File_Descriptor} in use, allowing us to write to 2 locations at once.
  \end{itemize}
\item[\cinline{int close(int fd)}] Close this \nameref{def:File_Descriptor}.
  \begin{itemize}
  \item Delete the \nameref{def:File_Descriptor} and deallocate the \nameref{def:File_Description}.
  \item Once all \nameref{def:File_Description}s of that point to this \nameref{def:vnode} are closed, the vnode is freed and the \nameref{def:File} is really closed.
  \end{itemize}

\item[\cinline{off_t lseek(int fd, off_t offset, int whence)}]
\item[\cinline{ssize_t read(int fd, void *buf, size_t nbytes)}]
  \begin{itemize}
  \item Read up to \texttt{nbytes} from \texttt{fd} into the buffer \texttt{buf}.
  \item Blocks until at least one byte is available.
  \item Returns the number of bytes \textbf{actually} read.
  \end{itemize}

\item[\cinline{ssize_t write(int fd, void *buf, size_t nbytes)}]
  \begin{itemize}
  \item Write up to \texttt{nbytes} into the open file at \texttt{fd} from \texttt{buf}.
  \item Returns the number of bytes \textbf{actually} written.
  \end{itemize}
\end{description}

\subsubsection{Why up to \texttt{nbytes}?}\label{subsubsec:Why_up_to_nbytes}
The \texttt{nbytes} parameters in \cinline{read} and \cinline{write} are necessary because the kernel attempts to maximize performance and minimize throughput.
By knowing these terms, it makes it easier to schedule disk reads and writes, and allows for more efficient kernel buffering.

Each of the items in the list below are each valid reasons to have to specify the number of bytes to manipulate.
\begin{itemize}[noitemsep]
\item \cinline{read}
  \begin{itemize}[noitemsep]
  \item \texttt{EOF}
  \item Unreadable \texttt{fd}, for example when the disk is failing.
  \item If the file is slow, this allows reading, a quick return to the \nameref{def:Process} ASAP and allow it to decide to read again or do something else.
  \item Interrupt, for example from a \nameref{def:Context_Switch}.
  \end{itemize}
\item \cinline{write}
  \begin{itemize}[noitemsep]
  \item Out of space, for example when the file is full, or when there is no more storage space on the backing media.
  \item Unwritable \texttt{fd}
  \item Slow File
  \item Interrupt
  \end{itemize}
\end{itemize}

\subsubsection{Non-Blocking Reads}\label{subsubsec:Non-Blocking_Reads}
Non-blocking reads are just like the other \texttt{read} operation, but it returns to the calling process immediately.

\subsection{Buffering}\label{subsec:Buffering}
\begin{definition}[Buffering]\label{def:Buffering}
  \emph{Buffering} means the system will read more bytes than we (the programmer) actually requested, and will place all of the gathered data into a separate \emph{backing buffer}.
\end{definition}

How it works:
\begin{enumerate}[noitemsep]
\item The user requests an amount of data
\item The kernel fetches it, plus some extra
\item The extra is not presented to the user, but is stored in the kernel
\item If the user asks for the next part, which is in the extra part, the read returns immediately.
\end{enumerate}

\subsubsection{Benefits}\label{subsubsec:Buffering_Benefits}
Your computer has many levels of buffers.
\begin{itemize}[noitemsep]
\item User-level buffers
\item Kernel-level buffers
\item There are also lower-level buffers/caches. They are present in:
  \begin{itemize}[noitemsep]
  \item The CPU
  \item The storage controller
  \item The disk itself
  \end{itemize}
\end{itemize}

By having many levels of buffers, we can create the illusion of great speeds, depending on the workload.
This way, we can operate on a buffer in-memory, rather than having to go to disk constantly.
That also has the added benefit of avoiding extra \nameref{def:Syscall}s.

\subsection{File I/O API}\label{subsec:File_IO_API}
\begin{description}[noitemsep]
\item[\cinline{FILE* fopen(const char *filename, const char *mode)}] Open the file specified by \texttt{filename} in the mode specified by \texttt{mode}.
\item[\cinline{FILE* fdopen(int fd, const char *mode)}] Open the file described by the \nameref{def:File_Descriptor} in the mode specified by \texttt{mode}.
\item[\cinline{int fclose(FILE *stream)}] Close the file (object) stream that \texttt{stream} points to.
\item[\cinline{int fseek(FILE *stream, long offset, int whence)}] Seek within the file stream specified by \texttt{stream}, according to the \texttt{offset}, from \texttt{whence}.
\item[\cinline{size_t fread(void *ptr, size_t size, size_t nitems, FILE *stream)}] Read \texttt{nitems}, each of size \texttt{size}, from \texttt{stream}, and store the contents where \texttt{ptr} points to.
\item[\cinline{size_t fwrite(void *ptr, size_t size, size_t nitems, FILE *stream)}] Write \texttt{nitems}, each of size \texttt{size}, from \texttt{ptr} to \texttt{stream}.
\item[\cinline{int fprintf(FILE *stream, const char *format, ...)}] Write a character stream specified by \texttt{format}, with the dots being additional parameters as specified by \texttt{format}, to \texttt{stream}.
\item[\cinline{int fscanf(FILE *stream, const char *format, ...)}] Read a character stream specified by \texttt{format}, with the dots being additional parameters as specified by \texttt{format}, to \texttt{stream}.
\item[\cinline{char* fgets(char *str, int size, FILE *stream)}] Get \texttt{size} characters from \texttt{stream} and store it where \texttt{str} points to.
\end{description}

\subsection{Generalizing}\label{subsec:Generalizing}
An example of a higher-level I/O function that performs many of the tasks we have discussed here is \texttt{printf}.
It handles the:
\begin{enumerate}[noitemsep]
\item \nameref{def:File_Descriptor}, \texttt{STDOUT}
  \begin{itemize}[noitemsep]
  \item Way to access the file in the user-level code.
  \end{itemize}
\item The Backing Buffer
  \begin{itemize}[noitemsep]
  \item Significantly reduces the cost of accessing stored information, because using memory access
  \end{itemize}
\item The number of unused bytes
\item Pointer to the next byte
\end{enumerate}

\subsubsection{Easing Your Life}\label{subsubsec:Easing_Your_Life}
While \texttt{open} and \texttt{read} are the lowest level of I/O \nameref{def:Syscall}, and are available to you, they are also somewhat hard to work with.
So, there are a variety of syscalls that wrap \texttt{open} for you, performing many of the actions we discussed earlier.
Some examples are:
\begin{itemize}[noitemsep]
\item \texttt{fclose}
\item \texttt{fflush}
\item \texttt{fgets}
\end{itemize}

All of these examples are from the system-library (which is really from \texttt{glibc}) \texttt{stdio.h}.

\paragraph{Stream Objects}\label{par:Stream_Objects}
The functions from \texttt{stdio.h} operate on \emph{stream} objects, which are wrappers around a file descriptor and the associated buffer.
This means stream buffers can absorb multiple writes before flushing to the underlying file.
However, a flush would occur when:
\begin{itemize}[noitemsep]
\item The buffer is full
\item Normal Process Termination
\item A newline is encountered.
  In a line-buffered stream, for example, printing to the screen.
\item Explicitly, with the \texttt{fflush} function.
\end{itemize}

However, this means that if the data is \textbf{not} flushed from the buffer, then the data is lost.
This could be a dangerous operation because the programmer might expect the data to be flushed/written immediately.

\subsection{Performing Both Input and Output}\label{subsec:Performing_Input_and_Output}
Performing both input and output at the same time to the same backing buffer is undefined int he C compiler, however, the C standard \textbf{does} specify this.
\begin{itemize}[noitemsep]
\item ISO C99 standard 7.19.5.3 Paragraph 6
\item Output shall not be followed by input without an intervening call to the \texttt{fflush} function.
\item Input shall not be directly followed by output without an intervening call to a file positioning function.
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../CS_351-Systems_Programming-Reference_Sheet"
%%% End:
