\subsection{Disk Management}\label{subsec:Disk_Management}
The OS handles many other aspects of the disks and their usage as well.

\subsubsection{Formatting}\label{subsubsec:Formatting}
There are 2 kinds of formatting:
\begin{enumerate}[noitemsep]
\item \nameref{def:Low_Level_Formatting}
\item \nameref{def:High_Level_Formatting}
\end{enumerate}

Most hard disks are low-level-formatted at the factory as a part of the manufacturing process.
This enables the manufacturer to test the disk and to initialize the mapping from logical block numbers to defect-free sectors on the disk.

\begin{definition}[Low-Level Formatting]\label{def:Low_Level_Formatting}
  \emph{Low-Level Formatting} or \emph{Physical Formatting} is the process of dividing the sectors in the tracks on the platters.
  This fills each sector with a special data structure typically consisting of a header, a data area (usually 512 bytes), and a trailer.
  The header and trailer contain information used by the disk controller: a sector number and an error-correcting code.
\end{definition}

The ECC contains enough information, if only a few bits of data have been corrupted, to enable the controller to identify which bits have changed and calculate what their correct values should be.
It then reports a recoverable soft error.
The controller automatically does the ECC processing whenever a sector is read or written.

It is usually possible to choose among a few data sizes, such as 256, 512, and 1,024 bytes.
Formatting a disk with a larger sector size means that fewer sectors can fit on each track; but it also means that fewer headers and trailers are written on each track and more space is available for user data.
\nameref{def:Operating_System}s may not support sizes other than a sector size of 512 bytes.

\begin{definition}[High-Level Formatting]\label{def:High_Level_Formatting}
  \emph{High-Level Formatting} or \emph{Logical Formatting} is the process an operating system goes through to partition the disk into groups of cylinders and put a file system onto these partitions.
  Each partition can be logically viewed as a separate disk, which can have its own file system.
  This stores the initial file system data structures, which can include maps of free and allocated storage and an initial empty directory.
\end{definition}

To increase efficiency, most \nameref{def:File_System}s group blocks together into larger chunks, frequently called clusters.
Disk I/O is done via blocks, but file system I/O is done via clusters.

Sometimes we want to use a disk partition just as a large sequential array of logical blocks, without any structure.
This array is called a raw disk, and I/O to this array is termed raw I/O.
Because there are no structures, and hence no file syste, raw I/O bypasses all file-system services, such as the buffer cache, file locking, prefetching, space allocation, file names, and directories.

\subsubsection{Boot Block}\label{subsubsec:Boot_Block}
A computer requires a bootstrap program that starts the computer when it is given a power-on command.
The first part of this involves starting the CPU from a cold, unpowered, state and initializing it and putting the \nameref{def:Kernel} into memory.

The bootstrap program lives in ROM, on the motherboard itself.
Because it is ROM, the CPU always knows where to find it, and the program stored there is protected from infection.
This program is just step one of 2 for booting, and is reponsible for finding the code for step 2.
It looks for and brings in the \nameref{def:Bootloader} from disk.
At this point, there are no device drivers, because those are provided by the \nameref{def:Kernel}.

The \nameref{def:Bootloader} can be changed easily, as it is on read-write storage.
It is stored in the boot blocks at a fixed location on the disk.
Its responsibility is to load the entire \nameref{def:Operating_System} from a non-fixed location on-disk.
A disk containing the boot blocks/boot partition is a boot disk or system disk.

\subsubsection{Bad Blocks}\label{subsubsec:Bad_Blocks}
Frequently, one or more sectors become defective.
Most disks even come from the factory with bad blocks.
Depending on the disk and controller in use, these blocks are handled in a variety of ways.

However, there are 2 common ways to handle this problem:
\begin{enumerate}[noitemsep]
\item Manually handling the blocks when they go bad.
  Any bad blocks that are discovered are flagged as unusable so that the file system does not allocate them.
  If blocks go bad during normal operation, a special program must be run manually to search for the bad blocks and to lock them away.
  Data that resided on the bad blocks usually is lost.

\item \textbf{Sector Sparing} or \textbf{Forwarding} has the disk manage these blocks itself.
  The controller maintains a list of bad blocks on the disk.
  The list is initialized during the low-level formatting at the factory and is updated over the life of the disk.
  Low-level formatting also sets aside spare sectors not visible to the operating system.
  The controller can be told to replace each bad sector logically with one of the spare sectors.
\end{enumerate}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
