\input{Reference_Sheet_Preamble}

% These packages are more specific to certain documents, but will be availabe in the template
% \usepackage{esint} % Provides us with more types of integral symbols to use
% \usepackage[outputdir=./TeX_Output]{minted} % Allow us to nicely typeset 300+ programming languages
% This document must be compiled with the -shell-escape flag if the packages above are uncommented

\graphicspath{{./Drawings/EITP20-Secure_Systems_Engineering}} % Uncomment this to use pictures in this document
% \addbibresource{./Bibliographies/CourseNum-Name.bib}

% Math Operators that are useful to abstract the written math away to one spot
% These are supposed to be document-specific mathematical operators that will make life easier
% Many fundamental operators are defined in Reference_Sheet_Preamble.tex

\begin{titlepage}
  \title{EITP20: Secure Systems Engineering --- Final Exam Study Questions}
  \author{Karl Hallsby}
  \date{Last Edited: \today} % We want to inform people when this document was last edited
\end{titlepage}

\begin{document}
\pagenumbering{gobble}
\maketitle
\pagenumbering{roman} % i, ii, iii on beginning pages, that don't have content
\tableofcontents
\clearpage
\pagenumbering{arabic} % 1,2,3 on content pages

\section{Lecture 1 --- Threat Analysis}
\begin{itemize}
\item Which are the steps involved in the overall design process of a secure system?
  \begin{itemize}[noitemsep]
  \item Describe the relationship between the different steps?
  \end{itemize}

\item In which way do a use-case description assist in the secure system design process?
\item What is a threat analysis?
\item What is the main purpose of performing a threat analysis?
\item Can you give example of threat analysis approaches?
\item What is the purpose of a security requirements list?
  \begin{itemize}[noitemsep]
  \item Can you list input sources for deriving security requirements?
  \item Which are the duties of the security engineering in the security requirements gathering process?
  \end{itemize}

\item Can you elaborate on the main differences between security requirements and other system requirements?
\item Please list at least three different “types” of security architectures?
  \begin{itemize}[noitemsep]
  \item Explain the main differences between the different
    listed types?
  \end{itemize}

\item Which are the main steps preceeding the actual secuirty design step?
\item Explain the main design choices that need to be done at the security design process.
\item List and explain different type of security evaluations that typically are done ”in house”
\item List and explain different type of security evaluations that typically are done by external experts
\item What is a pen test and what is the purpose of such test?
\item What is a protocol analysis tool and what is the purpose of using such tool?
\item What is the Common Criteria (CC) standard?
  \begin{itemize}[noitemsep]
  \item List the 7 evaluation levels defined in CC and explain the main differences between the levels?
  \item List the three different system documents part of a CC and describe their main purpose.
  \end{itemize}
\item What is CISSP?\@
\end{itemize}

\end{document}