\documentclass[10pt,letterpaper,final,twoside,notitlepage]{article}
\usepackage[margin=.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm} % Gives us plain, definition, and remark to use in \theoremstyle{style}
\usepackage{graphicx}

\usepackage{hyperref} % Generate hyperlinks to referenced items
\usepackage{nameref} % Can make references by name to places
\usepackage{ctable} % Greater control over tables and how they look
\usepackage{subcaption} % Allows for multiple figures in one Figure environment
\usepackage{textcomp} % Forcibly loads a package that gensymb relies on
\usepackage{gensymb} % Gives access to some characters, for example, the degree symbol
\usepackage{enumitem} % Provides [noitemsep, nolistsep] for more compact lists
\usepackage{chngcntr} % Allows us to tamper with the counter a little more
\usepackage{empheq} % Allow boxing of equations in special math environments

%\graphicspath{{./Drawings/Class_Name_and_Number}} % Uncomment this to use pictures in this document
%\numberwithin{equation}{section} % Uncomment this to number equations with section numbers too

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem{definition}{Defn}
\newtheorem{corollary}{Corollary}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}[definition]
\newtheorem*{remark*}{Remark}
%\counterwithin{definition}{subsection} % Uncomment to have definitions use section.subsection numbering

\newcounter{example}[section]
\newenvironment{example}[1] % Define an example environment
	{\refstepcounter{example} \begin{center}\begin{tabular}{|p{0.9\textwidth}|} \hline \textbf{Example~\theexample.} \textit{#1}  \\ }
	{\\ \hline \end{tabular} \end{center}}

\renewcommand\qedsymbol{$\blacksquare$} % Change proofs to have black square at end

\DeclareMathOperator{\RealNums}{\mathbb{R}}

\author{Karl Hallsby}
\title{Math 251 Differential Equations}

\begin{document}
\section{Introduction} \label{sec:Introduction}
	\begin{example}{Confirm Differential Solution}
		Test example.
	\end{example}
	\subsection{Definitions and Terminology} \label{subsec:Definitions and Terminology}
		\begin{definition}[Differential Equation] \label{def:Differential Equation}
			A \emph{differential equation (DE)} is an equation with 1 or more derivatives.
			\begin{remark}
				The highest differential determines the order of the differential equation.
				This means that the differential equation below is of order 2.
				\begin{align*} 
					y'' + y &= 0 \\
					\frac{d^{2}y}{dx^{2}} + y &= 0 \\
				\end{align*} 
			\end{remark}
		\end{definition}
		\begin{definition}{Initial Value Problem} \label{def:Initial Value Problem}
			A differential equation with one or more initial conditions is called an \emph{initial value problem (IVP)}.
			\begin{remark}
				To solve an initial value problem, you must have the same number of initial conditions as the order of the differential equation.
			\end{remark}
			\begin{remark}[Existence of Unique Solution]
				$R$ is a rectangular region on the xy-plane $a \leq x \leq b$, $c \leq y \leq d$ that contains $\left( x_{0}, y_{0} \right)$ interior.
				If $f \left( x,y \right)$ and $\frac{df}{dy}$ are continuous on $R$, then an interval exists $I_{0}$ such that $\left( x_{0}-h, x_{0}+h \right)$ where $h>0$, on the interval $\left[ a,b \right]$, and a unique function $y \left( x \right)$, defined on $I_{0}$ that is a solution of the initial value problem.
			\end{remark}
		\end{definition}
	
	\subsection{Confirm If Differential Equation} \label{subsec:Confirm Differential Equation}
	You can confirm if the solution $y \left( x \right)$ found for a differential equation $y \left( x \right)'$ is the solution by differentiating the solution and putting that in the solved differential equation and verfiying that the equation holds true.
	This is shown in Example~\ref{ex:Confirm Differential Solution}
	\begin{example}{} \label{ex:Confirm Differential Solution}
		Given the differential equation, $2y' + y = 0$, is $y = e^{\frac{-x}{2}}$ a solution? \newline
		\begin{align*}
			y'% &= \frac{-1}{2} e^{\frac{-x}{2}} \\
%			2 \left( \frac{-1}{2} e^{\frac{-x}{2}} \right) + \left( e^{\frac{-x}{2}} \right) &= 0 \\
%			-e^{\frac{-x}{2}} + e^{\frac{-x}{2}} &= 0 \\
%			0 &= 0 \text{ \checkmark}
		\end{align*}
	\end{example}
		
	\subsection{Separable Differential Equation} \label{subsec:Separable Differential Equation}
		\begin{definition}[Separable] \label{def:Separable}
			A \emph{separable} differential equation allows you to move various elements around to solve the equation.
			For example,
			\begin{align*}
				\frac{dP}{dt} &= kP \\
				\frac{1}{P} dP &= k dt \\
				\ln \left( P \right) &= kt + C \\
				P &= Ce^{kt}
			\end{align*}
			\begin{remark}
				These are used extensively in modelling phenomena with differential equations.
				These include: \nameref{subsubsec:Population Growth}, \nameref{subsubsec:Radioactive Decay}, \nameref{subsubsec:Newton Law of Cooling/Heating}, and \nameref{subsubsec:Spread of Disease}.
			\end{remark}
		\end{definition}
		
	\subsection{Modeling with Differential Equations} \label{subsec:Modeling with DEs}
		\subsubsection{Population Growth} \label{subsubsec:Population Growth}
			\begin{definition}[Population Growth]
				\emph{Population growth} can be modelled with a separable differential equation. Namely,
				\begin{equation} \label{eq:Population Growth}
					\frac{dP}{dt} = kP
				\end{equation}
				\begin{remark}[Population Growth Parameters] \label{rmk:Population Growth Parameters}
					The parameters for the \nameref{eq:Population Growth}~equation are given below.
					\begin{itemize}[noitemsep, nolistsep]
						\item $k>0$
						\item $P>0$
					\end{itemize}
				\end{remark}
			\end{definition}
		\subsubsection{Radioactive Decay} \label{subsubsec:Radioactive Decay}
			\begin{definition}[Radioactive Decay] \label{def:Radioactive Decay}
				\emph{Radioactive decay} is the process that some particularly heave atoms undergo.
				\begin{definition}[Half-Life] \label{subdef:Half-Life}
					The \emph{half-life} is the usual reported metric, and is defined as the amount of time required for an element to half its mass through \nameref{def:Radioactive Decay}.
				\end{definition}
				\begin{equation} \label{eq:Radioactive Decay}
					\frac{1}{2} A_{0} = A_{0} e^{kt}
				\end{equation}
				\begin{remark}[Radioactive Decay Parameters] \label{rmk:Radioactive Decay Parameters}
					The parameters for the \nameref{eq:Radioactive Decay}~equation are given below.
					\begin{itemize}[noitemsep, nolistsep]
						\item $k<0$
						\item $A>0$
					\end{itemize}
				\end{remark}
			\end{definition}
		\subsubsection{Newton's Law of Cooling/Heating} \label{subsubsec:Newton Law of Cooling/Heating}
			\begin{definition}[Newton's Law of Cooling/Heating] \label{def:Newton Law of Cooling/Heating}
				\emph{Newton's Law of Cooling/Heating} is the same equation, but some of the parameters change.
				This equation is defined as:
				\begin{equation} \label{eq:Newton Law of Cooling/Heating}
					\frac{dT}{dt} = k \left( T-T_{m} \right)
				\end{equation}
				\begin{remark} \label{rmk:Newton Law of Cooling/Heating Parameters}
					The parameters for the \nameref{eq:Newton Law of Cooling/Heating}~equation are given below.
					\begin{itemize}[noitemsep, nolistsep]
						\item $\frac{dT}{dt}$; The rate of change of temperature in the object per unit time.
						\item $k<0$; The cooling constant and is unique to every object.
						\item $T$; The starting temperature.
						\item $T_{m}$; The temperature of the surrounding medium.
					\end{itemize}
				\end{remark}
			\end{definition}
		\subsubsection{Spread of Disease} \label{subsubsec:Spread of Disease}
			\begin{definition}[Spread of Disease] \label{def:Spread of Disease}
				This is used to model the spread of something throughout a society or group of people.
				\begin{equation} \label{eq:Spread of Disease}
					\frac{dx}{dt} = kxy
				\end{equation}
				\begin{remark}
					The parameters for the \nameref{eq:Spread of Disease}~equation are given below.
					\begin{itemize}[noitemsep, nolistsep]
						\item $\frac{dx}{dt}$; Change in the number of infected per unit time.
						\item $k<0$; Transmission Constant
						\item$x$; Number of Infected
						\item $y$; Number of non-infected, $y$ is really a function of $x$
						\begin{itemize}[noitemsep, nolistsep]
							\item $y = n+1-x$
						\end{itemize}
					\end{itemize}
				\end{remark}
			\end{definition}
		
		\subsubsection{Chemical Reactions} \label{subsubsec:Chem Reactions}
		\subsubsection{Tank Mixture} \label{subsubsec:Tank Mixture}
		\subsubsection{Torricelli's Law} \label{subsubsec:Torricelli's Law}
		\subsubsection{LRC Circuits} \label{subsubsec:LRC Circuits}
			\begin{definition}[LRC Circuits] \label{def:LRC Circuits}
				An \emph{LRC Circuit} is analyzed in terms of the energy moving through the circuit.
				There is a unique relationship for the energy in each element:
				\begin{equation} \label{eq:Energy in Capacitor}
					E \left( t \right) = \frac{q}{C}
				\end{equation}
				\begin{equation} \label{eq:Energy in Resistor}
					E \left( t \right) = RI = R \frac{dq}{dt}
				\end{equation}
				\begin{equation} \label{eq:Energy in Inductor}
					E \left( t \right) = L \frac{dI}{dt} = L \frac{d^{2}q}{dt^{2}}
				\end{equation}
				\begin{remark}
					Depending on the circuit given, you might use a combination of these, but you \emph{\textbf{must}} have at least one capacitor or inductor, otherwise it is not a differential equation.
				\end{remark}
				\begin{remark}
					These equations \emph{add} together when the entire circuit is in series, i.e. the elements are put together back-to-back.
				\end{remark}
			\end{definition}
	
	\subsection{Linear and Non-Linear Differential Equations} \label{Linear vs. Non-Linear Differential Equations}
		\begin{definition}[Linear Differential Equation] \label{def:Linear Differential Equation}
			A \emph{linear differential equation} is one that satisfies one of the following equations below.
			\begin{equation} \label{eq:Linear Differential Equation}
				\begin{aligned}
					a_{1} \left( x \right) \frac{dy}{dx} + a_{0} \left( x \right) &= g \left( x \right) \\
					a_{2} \left( x \right) \frac{d^{2}y}{dx^{2}} + a_{1} \left( x \right) \frac{dy}{dx} + a_{0} \left( x \right) &= g \left( x \right) \\
				\end{aligned}
			\end{equation}
			\begin{remark}
				The equations in Equation~\eqref{eq:Linear Differential Equation} can be generalized to the $n$th order as shown below.
				\begin{equation} \label{eq:General Linear Differential Equation}
					a_{n} \left( x \right) \frac{d^{n}y}{dx^{n}} + a_{n-1} \left( x \right) \frac{d^{n-1}y}{dx^{n-1}} + \ldots + a_{1} \left( x \right) \frac{dy}{dx} + a_{0} \left( x \right) = g \left( x \right)
				\end{equation}
			\end{remark}
		\end{definition}
		\begin{definition}[Non-Linear] \label{def:Non-Linear Differential Equation}
			A \emph{non-linear} differential equation is one that does not satisfy the definition of a \nameref{def:Linear Differential Equation}.
			It does not obey Equation~\eqref{eq:General Linear Differential Equation}.
		\end{definition}

%====================================APPENDIX====================================
\clearpage
\appendix
\counterwithin{equation}{section}

\section{Reference Material} \label{sec:Reference Material}
\input{./Appendix-Reference_Material/Trig}
\input{./Appendix-Reference_Material/Calculus}
\end{document}