\subsection{Mechanisms of Energy Transfer}\label{subsec:Mechanisms_Energy_Transfer}
From our discussion of changes in energy in \Cref{subsec:Divisions_of_Energy}, we know that there are 3 main mechanisms of energy transfer:
\begin{enumerate}[noitemsep]
\item \nameref{def:Heat_Transfer}
\item \nameref{def:Work}
\item Mass Flow
\end{enumerate}

In addition, according to the \nameref{def:Law_Conservation_Energy}, we know that the total amount of energy must be conserved throughout a process.
This leads to 2 separate types of processes occurring.
\begin{description}[noitemsep]
\item[Cyclical] A cyclic process is one where the beginning and end states are identical.
  This also means that $\Heat_{In} = \Work_{Out}$ and $\Heat_{Out} = \Work_{In}$, making $\Change{\Energy} = 0$, and $\Energy_{In} = \Energy_{Out}$.
\item[Regular] A regular process is the general way to deal with processes, and cyclic processes are ones that are special cases of a regular process.
  In this case, there is a change in energy $\Change{\Energy} = (\Heat_{In} - \Heat_{Out}) + (\Work_{In} - \Work_{Out})$.
\end{description}

Now, we can generalize \Cref{eq:Change_Total_Energy} even more, to allow for the changing of systems.
We start with an equation that represents the energy of the entire \nameref{def:System}, \Cref{eq:System_Energy}.

\begin{equation}\label{eq:System_Energy}
  \InternalEnergy + \KineticEnergy + \PotentialEnergy + \FlowEnergy = \Heat + \Work
\end{equation}

Now, if we want to discuss the change in the total energy of the system, we can use \Cref{eq:Change_System_Energy}.

\begin{equation}\label{eq:Change_System_Energy}
  \begin{aligned}
    \Change{\InternalEnergy} + \Change{\KineticEnergy} + \Change{\PotentialEnergy} + \Change{\FlowEnergy} &= \Change{\Heat} + \Change{\Work} \\
    &= (\Heat_{In} - \Heat_{Out}) + (\Work_{In} - \Work_{Out})
  \end{aligned}
\end{equation}

The implications of \Cref{eq:Change_System_Energy} are listed below.
\begin{itemize}[noitemsep]
\item For stationary systems, $\Change{\KineticEnergy} = 0$, and $\Change{\PotentialEnergy} = 0$.
\item For stationary and \nameref{def:Closed_System}s (when there is \textbf{no} mass transfer) $\Change{\KineticEnergy} = 0$, $\Change{\PotentialEnergy} = 0$, $\Change{\FlowEnergy} = 0$.
  Therefore, $\Change{\InternalEnergy} = \Change{\Heat} + \Change{\Work}$.
\end{itemize}

\begin{example}[Textbook Problem 5.29]{Enthalpy of Saturated Water}
  Saturated water vapor at $\Temp = \SI{200}{\degreeCelsius}$ and is \nameref{def:Isothermal}ly condensed into a \nameref{def:Saturated_Liquid} in a piston-cylinder device.
  Calculate the \nameref{def:Heat_Transfer} and \nameref{def:Work} done during this process in \si{\kilo\joule\per\kilo\gram}?
  \tcblower{}
  \textbf{Concepts:} \\
  A piston-cylinder means that the pressure is constant throughout the duration of a \nameref{def:Process}. \\
  An \nameref{def:Isothermal} process is one in which the temperature is constant. \\
  The water is moving from a \nameref{def:Saturated_Vapor} to a \nameref{def:Saturated_Liquid}. \\
  The total energy is conserved throughout the \nameref{def:Process}

  \textbf{Explore:} \\
  From our understanding of \nameref{def:Saturated_Mixture}s, we know that we should use Table A.4 from the textbook's appendix.
  This is both a stationary system and a \nameref{def:Closed_System}, so $\Change{\KineticEnergy} = 0$, $\Change{\PotentialEnergy} = 0$, and $\Change{\FlowEnergy} = 0$.
  Therefore, the only terms in \Cref{eq:Change_System_Energy} we need to care about are
  \begin{equation*}
    \Change{\InternalEnergy} = \Change{\Heat} + \Change{\Work}
  \end{equation*}

  To simplify this further, we can assume that at the starting state, there was no \nameref{def:Heat} going in and no \nameref{def:Work} being done.
  Therefore, $\Heat_{In} = 0$ and $\Work_{In} = 0$.
  So, the equation above simplifes down to
  \begin{align*}
    \InternalEnergy_{2} - \InternalEnergy_{1} &= -\Heat_{Out} + -\Work_{Out} \\
                                              &= -\Heat_{Out} - \Pressure(\Volume_{2} - \Volume_{1})
  \end{align*}

  The difference in \nameref{def:Internal_Energy} is actually recorded in the textbook's Table A.4 as $\InternalEnergy_{fg}$.

  \textbf{Plan:} \\
  Solve for the above equation, using Table A.4 in the textbook to find the value of the change in \nameref{def:Internal_Energy} of the system.

  \textbf{Solve:} \\
  From Table A.4, we know that at $\Temp = \SI{200}{\degreeCelsius}$, $\SaturatedPressure = \SI{1555}{\kilo\pascal}$.
  We also know $\Fluid{\SpecificVolume} = \SI{0.001157}{\meter\cubed\per\kilo\gram}$ and $\Vapor{\SpecificVolume} = \SI{0.12721}{\meter\cubed\per\kilo\gram}$.
  Lastly, we can also find $\Vapor{\SpecificInternalEnergy} = \SI{2594}{\kilo\joule\per\kilo\gram}$ and $\Fluid{\SpecificInternalEnergy} = \SI{850.46}{\kilo\joule\per\kilo\gram}$.

  \begin{align*}
    \Work_{Out} &= 1555 (0.001157 - 0.12621) \\
                &= \SI{-196}{\kilo\pascal\meter\cubed\per\kilo\gram} \\
                &= \SI{-196}{\kilo\joule\per\kilo\gram} \\
    \Fluid{\SpecificInternalEnergy} - \Vapor{\SpecificInternalEnergy} &= 850 - 2594 \\
                &= \SI{-1744}{\kilo\joule\per\kilo\gram} \\
    \Fluid{\SpecificInternalEnergy} - \Vapor{\SpecificInternalEnergy} &= -\Heat_{Out} - \Work_{Out} \\
    \SI{-1744}{\kilo\joule\per\kilo\gram} &= -\Heat_{Out} - \SI{-196}{\kilo\joule\per\kilo\gram} \\
    \Heat_{Out} &= \SI{1940}{\kilo\joule\per\kilo\gram}
  \end{align*}

  Thus, the vapor is cooling and condensing in the piston-cylinder, which makes sense in this case.

  \textbf{Validate:} \\
  We can validate this by performing the same calculations using the change in \nameref{def:Specific_Enthalpy} instead.
  \begin{align*}
    \Change{\SpecificEnthalpy} &= -\Heat_{Out} \\
    \Fluid{\SpecificEnthalpy} - \Vapor{\SpecificEnthalpy} &= -\Heat_{Out} \\
    \intertext{The change in specific enthalpy can be found in Table A.4 again.}
    (852.3 - 2792) \si{\kilo\joule\per\kilo\gram} &= -\Heat_{Out} \\
    \Heat_{Out} &= \SI{1939.7}{\kilo\joule\per\kilo\gram} \\
    \shortintertext{Now, taking significant figures.}
    \Heat_{Out} &= \SI{1940}{\kilo\joule\per\kilo\gram}
  \end{align*}

  \textbf{Generalize:} \\
  Any time the system is changing states, we can use \nameref{def:Enthalpy} and/or \nameref{def:Specific_Enthalpy}.
  This helps us avoid these longer calculations using all the other forms of \nameref{def:Energy} in the \nameref{def:System}.
\end{example}

\begin{example}[Textbook Problem 5.34]{Case when Enthalpy Cannot be Used}
  Rigid $\Volume = \SI{1}{\feet\cubed}$ vessel that contains R-134a.
  The refrigerant is at $\Temp_{1} = \SI{-20}{\degreeF}$ with \nameref{def:Quality} $\SI{27.7}{\percent}$.
  It is heated until it reaches $\Temp_{2} = \SI{100}{\degreeF}$.
  Calculate the \nameref{def:Heat_Transfer} required to do this?
  \tcblower{}
  \textbf{Concepts:} \\
  Because we are told the \nameref{def:Quality}, we know this is a \nameref{def:Saturated_Mixture}. \\
  This is a rigid vessel, so the volume is constant, implying this is both a stationary and closed system. \\
  Because $\Heat$ is an \nameref{def:Energy}, not \nameref{def:Specific_Energy}, we need to keep it as an \nameref{def:Extensive_Property}.
  Using all of these facts, $\Change{\InternalEnergy} = \Change{\Heat} + \Change{\Work}$.

  \textbf{Explore:} \\
  We can start by looking at the various components of the equation above.
  There is no cooling going on in this system, so $\Heat_{Out} = 0$.
  There is no work being done on the refrigerant, so both $\Work_{In} = 0$ and $\Work_{Out} = 0$.
  Therefore, $\Change{\InternalEnergy} = \Heat_{In}$, which makes sense, as we were told we were only heating the refrigerant.
  Lastly, because the refrigerant starts as a \nameref{def:Saturated_Mixture}, we need to use:
  \begin{align*}
    \SpecificVolume_{x} &= \Fluid{\SpecificVolume} + \Quality (\Vapor{\SpecificVolume} - \Fluid{\SpecificVolume}) \\
    \SpecificInternalEnergy_{x} &= \Fluid{\InternalEnergy} + \Quality (\Vapor{\InternalEnergy} - \Vapor{\InternalEnergy})
  \end{align*}

  \textbf{Plan:} \\
  Solve $\Change{\SpecificInternalEnergy} = \Heat_{In}$ by finding the \nameref{def:Internal_Energy} values in the English appendix table for R-134a.

  \textbf{Solve:} \\
  When $\Temp = \SI{-20}{\degreeF}$, $\SaturatedPressure = \SI{12.9}{\psia}$.
  In addition, the \nameref{def:Quality} is $\Quality = 0.227$.
  Because this is in a rigid vessel, the volume is constant, which also means the \nameref{def:Specific_Volume} is constant.
  \begin{align*}
    \SpecificVolume_{x} &= \Fluid{\SpecificVolume} + \Quality (\Vapor{\SpecificVolume} - \Fluid{\SpecificVolume}) \\
    \intertext{Grabbing $\Fluid{\SpecificVolume}$ and $\Vapor{\SpecificVolume}$ from the Table in the textbook's appendix.}
                        &= 0.01156 + 0.277 (3.4424 - 0.01156) \\
                        &= \SI{0.961843}{\feet\cubed\per\lbm}
  \end{align*}

  Now we focus on finding the \nameref{def:Internal_Energy} of the \nameref{def:Saturated_Mixture} at the given \nameref{def:Quality}.
  \begin{align*}
    \SpecificInternalEnergy_{x} &= \Fluid{\SpecificInternalEnergy} + \Quality (\Vapor{\SpecificInternalEnergy} - \Fluid{\SpecificInternalEnergy}) \\
                        &= 6.014 + 0.277 (91.90 - 6.014) \\
                        &= \SI{29.8044}{\btu\per\lbm}
  \end{align*}

  Now, we focus on the end state.
  Keep in mind that the \nameref{def:Specific_Volume} is constant.
  Looking in the \nameref{def:Saturated_Mixture} tables for this temperature shows us the refrigerant is actually a \nameref{def:Superheated_Vapor} now.
  When looking at Table A.13E, we notice that the \nameref{def:Specific_Volume} does not have a single entry in the table.
  So, we must \nameref{def:Interpolate} instead.
  \begin{align*}
    y &= y_{0} + (x_{1} - x) \frac{y_{1} - y_{0}}{x_{1} - x_{0}} \\
    \shortintertext{\nameref{def:Interpolate} the \nameref{def:Pressure}.}
    \Pressure &= \Pressure_{0} + (\SpecificVolume - \SpecificVolume_{1}) \frac{\Pressure_{1} - \Pressure_{0}}{\SpecificVolume_{1} - \SpecificVolume_{0}} \\
      &= \SI{50}{\psia} + (0.961843 - .9072) \frac{60 - 50}{1.1043 - 0.9072} \\
      &= \SI{52.7}{\psia} \\
    \shortintertext{Now, interpolating the \nameref{def:Internal_Energy}.}
    \SpecificInternalEnergy_{2} &= \SpecificInternalEnergy_{1} + (\SpecificVolume - \SpecificVolume_{1}) \frac{\SpecificInternalEnergy_{1} - \SpecificInternalEnergy_{0}}{\SpecificVolume_{1} - \SpecificVolume_{0}} \\
      &= \SI{111.16}{\btu\per\lbm} + (0.961843 - 0.9072) \frac{111.17 - 111.56}{0.90762 - 1.1043} \\
      &= \SI{111.3}{\btu\per\lbm}
  \end{align*}

  Now that we have the beginning and end values of the pressure and internal energy, we can calculate the change in internal energy.
  \begin{align*}
    \Change{\SpecificInternalEnergy} &= \SpecificInternalEnergy_{1} - \SpecificInternalEnergy_{x} \\
                             &= 111.3 - 29.8044 \\
                             &= \SI{81.49}{\btu\per\lbm}
  \end{align*}

  However, because the change we actually got was the change in specific \nameref{def:Internal_Energy}, we must multiply it by the mass of the refrigerant in the system.
  We know the \nameref{def:Specific_Volume} and the volume of the vessel, so we can find the mass.
  \begin{align*}
    \SpecificVolume &= \frac{\Volume}{\Mass} \\
    \Mass &= \frac{\Volume}{\SpecificVolume} \\
                    &= \frac{1}{0.961843} \\
    &= \SI{1.03967}{\lbm}
  \end{align*}

  Now, multiplying $\Change{\SpecificInternalEnergy}$ with the mass, we get
  \begin{align*}
    \InternalEnergy &= \Mass \SpecificInternalEnergy \\
                    &= 1.03967 \times 81.49 \\
                    &= 84.7228 \\
                    &= \SI{84.73}{\btu}
  \end{align*}
\end{example}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../MMAE_320-Thermo-Reference_Sheet"
%%% End:
