\section{Quantum Mechanics} \label{sec:Quantum Mech}
Quantum mechanics may seem strange or weird, but it really isn't.
	\begin{definition}[Hamiltonian] \label{def:Hamiltonian}
		The \emph{hamiltonian} is a function that contains all knowable information about a particle or system in classical mechanics.
		\begin{equation} \label{eq:Hamiltonian}
			\text{Hamiltonian} = K+U
		\end{equation}
		\begin{itemize}[noitemsep, nolistsep]
			\item $K$ is the Kinetic Energy of the particle of system of particles.
			\begin{itemize}[noitemsep, nolistsep]
				\item $K = \frac{1}{2} mv^{2}$
				\item This contains the velocity of the system, which includes direction.
			\end{itemize}
			\item $U$ is the Potential Energy of the particle or system of particles.
			\begin{itemize}[noitemsep, nolistsep]
				\item $U = mg \left( y_{2} - y_{1} \right)$
				\item Contains the force of the system, its location, etc.
			\end{itemize}
		\end{itemize}
	\end{definition}

	\subsection{Schr\"{o}dinger's Equation} \label{subsec:Schrodingers Equation}
		\begin{definition}[de Broglie Wavelength] \label{def:de Broglie Wavelength}
			The \emph{de Broglie wavelength, $\lambda$}, is the wavelength that is \emph{associated} with a particle.
			In other words, a particle does \emph{\textbf{not}} have a wavelength, but the equations describing the particle's actions have a wavelength.
			\begin{note} \label{note:de Broglie Wavelength Momentum}
				As a direction consequence of the \nameref{def:de Broglie Wavelength},
				\begin{equation}
					p = \frac{h}{\lambda}
				\end{equation}
				\begin{itemize}[noitemsep, nolistsep]
					\item $h$ is Planck's Constant
					\item $\lambda$ is the \nameref{def:de Broglie Wavelength}
				\end{itemize}
			\end{note}
		\end{definition}
		\begin{definition}[Schr\"{o}dinger's Equation] \label{def:Schrodinger's Equation}
			\emph{Schr\"{o}dinger's equation} is a 3-dimensional Cartesian, time-independent way to express the equivalent of a \nameref{def:Hamiltonian} for quantum effects, namely in terms of waves (Denoted with $\Psi$ or $\psi$).
			There are 2 equivalent ways to express \nameref{def:Schrodinger's Equation}.
			The first way, Equation~\eqref{eq:Schrodinger's Equation, Form 1}, is the most general form.
			The second way, Equation~\eqref{eq:Schrodinger's Equation, Form 2}, has expanded the nabla operator, $\nabla$.
			\begin{equation} \label{eq:Schrodinger's Equation, Form 1}
				\frac{- \hbar^{2}}{2m} \cdot \nabla^{2} \Psi + V\Psi = E\Psi
			\end{equation}
			\begin{equation} \label{eq:Schrodinger's Equation, Form 2}
				\frac{-\hbar}{2m} \left( \frac{\partial^{2} \Psi}{\partial x^{2}} + \frac{\partial^{2} \Psi}{\partial y^{2}} + \frac{\partial^{2} \Psi}{\partial z^{2}} \right) + V\Psi = E\Psi
			\end{equation}
			\begin{itemize}[noitemsep, nolistsep]
				\item $E$ is the \emph{\textbf{total}} energy of the system
				\item $V$ is the \emph{\textbf{potential energy}} of the system. It may be:
				\begin{enumerate}[noitemsep, nolistsep]
					\item Constant
					\item A function of time and/or space
				\end{enumerate}
				\item $\Psi$ is a solution to \nameref{def:Schrodinger's Equation}, and contains all that can be known about the system
				\item $\frac{-\hbar^{2}}{2m} \cdot \nabla\Psi$ is the \emph{\textbf{kinetic energy}} operator
				\begin{itemize}[noitemsep, nolistsep]
					\item $\nabla^{2} = \frac{\partial^{2}}{\partial x^{2}} + \frac{\partial^{2}}{\partial y^{2}} + \frac{\partial^{2}}{\partial z^{2}}$ is in the Cartesian coordinate system ($x$, $y$, $z$), and is called ``del squared'', or ``2nd-degree nabla.''
				\end{itemize}
				\item For each observable system, there exists a corresponding mathematical operator.
			\end{itemize}
			\begin{note} \label{note:Psi Probability Distribution Function of Particle Location}
				The probability distribution function of the position of a particle in the system can be found according to Equation~\eqref{eq:Psi Probability Distribution Function of Particle Location}.
				This is the multiplication of $\Psi$ with its complex conjugate, $\Psi^{*}$.
				\begin{equation} \label{eq:Psi Probability Distribution Function of Particle Location}
					\text{PDF} = \Psi \cdot \Psi^{*}
				\end{equation}
			\end{note}
		\end{definition}
		\subsubsection{Properties of Schr\"{o}dinger's Equation} \label{subsubsec:Properties of Schrodingers Equation}
			Solutions, $\Psi$, to \nameref{def:Schrodinger's Equation} have the following characteristics:
			\begin{enumerate}[noitemsep, nolistsep]
				\item They are single-valued
				\item They are piecewise continuous values
				\item $\frac{\partial \Psi}{\partial x}$, $\frac{\partial \Psi}{\partial y}$, $\frac{\partial \Psi}{\partial z}$ are all continuous
				\item $\iiint \Psi \Psi^{*} \partial x \partial y \partial z$ is square integrable, or $\iiint \Psi \Psi^{*} \partial x \partial y \partial z \neq \pm \infty$
			\end{enumerate}

		\input{./Phys_224-Sections/Proofs/Rough_Justification_Schrodingers_Equation}
	
		{\Large \emph{\textbf{This means that the 3-Dimensional problem is broken down into 3 1-Dimensional problems!}} Each dimension can be approached exactly the same way as above.}
	
	\subsection{Operators} \label{subsec:Operators}
		\begin{definition}[Operator] \label{def:Operators}
			In quantum mechanics, for each observable dynamical variable, there exists a corresponding \emph{operator}, which, if applied to the solution of the \nameref{eq:Schrodinger's Equation, Form 1}, will provide a value for that dynamical variable.
			\nameref{def:Operators} do something to the function immediately after the operator.
			An operator is nothing other than shorthand for instruction to do something to a function or number. For example, some common operators are:
			\begin{enumerate}[noitemsep, nolistsep]
				\item $+$
				\item $-$
				\item $\times$
				\item $\divideontimes$
				\item $\frac{\partial^{2}}{\partial t^{2}}$
				\item $\int$
				\item $\sum$
				\item $\nabla^{2}$
			\end{enumerate}
			Operators have some vector-like properties, but they are not vectors.
			Further, they do not always always \emph{commute}.
			Part of the main confusion arising from the public press about quantum mechanics is as statement like ``\textellipsis $A \times B$ is equal to $B \times A$.''
			This is simply sensationalizing the fact that some operators have commutative properties.
			One example is:
			\begin{equation*}
				y \times \frac{\partial f \left( y \right)}{\partial y} \neq \frac{\partial \left[ x \times f \left( y \right) \right]}{\partial y}
			\end{equation*}
			For each classical observable, there exists a corresponding quantum mechanical operator.
			Some examples of simple quantum mechanical operators are:
			\begin{itemize}[noitemsep, nolistsep] \label{item:List of Quantum Mechanical Operators}
				\item Kinetic Energy: $\frac{-\hbar^{2}}{2m} \nabla^{2}$
				\item Momentum: $\frac{-\hbar}{i} \frac{\partial}{\partial x}$
				\item Potential Energy: $V$ is simply multiplication by $V$
				\item Angular Momentum: $L_{z} = \mathbf{r} \times \frac{-\hbar}{i} \nabla^{2}$
			\end{itemize}
			Further, some, but not all, operators are linear.
			Fortunately, the quantum mechanics kinetic energy operators, $\nabla^{2} = \frac{\partial^{2}}{\partial x^{2}} + \frac{\partial^{2}}{\partial x^{2}} + \frac{\partial^{2}}{\partial x^{2}}$ is a linear operator.
			This allows us to \emph{reduce a 3-dimensional problem into 3 1-dimensional problems.}
		\end{definition}
	
	\subsection{Eigenvalues} \label{subsec:Eigenvalues}
		\begin{definition}[Eigenvalue] \label{def:Eigenvalues}
			Let $\widehat{\mathbf{O}}$ be some quantum mechanical operator.
			If $O$ and $\Psi$ are \emph{eigenvalues} and eigenfunctions of $\widehat{\mathbf{O}} \Psi = o \Psi$, the result $o$ will always be a precise number.
			\emph{Eigenvalues} represent the value of precise, exact values of the dynamical variables corresponding to the eigenfunction of the operator.
		\end{definition}
	The \emph{\textbf{easy}} way to find the value of a dynamical variable $o$, given the corresponding operator, $\widehat{\mathbf{O}}$ is as follows.
		\input{./Phys_224-Sections/Proofs/Find_Eigenvalue}
	
	\subsection{Expectation Values} \label{subsec:Expectation Values}
		\begin{definition}[Expectation Values] \label{def:Expectation Values}
			If $o$ and $\Psi$ are \textbf{not} eigenvalues and eigenfunctions of $\widehat{\mathbf{O}}$, e.g. $\widehat{\mathbf{O}} \Psi = z \phi$, where $z$ is some number, and $\phi$ is some function not equal to $\Psi$, the result $z$ will be the average of many possible observations. \newline
			This is usually written as $\langle z \rangle$ and is called an \emph{expectation value}. The expectation value is the expected average value of a number of measurements for a dynamical variable described by the operator $\widehat{\mathbf{O}}$.
		\end{definition}
	
	\subsection{Determine Dynamical Variables in Quantum Mechanics} \label{subsec:Determine Dynamical Variables in Quantum Mechanics}
		This section has an example where we solve for dynamical variables.
		\input{./Phys_224-Sections/Examples/Determine_Dynamical_Variables}