\subsection{Allocating Kernel Memory}\label{subsec:Allocating_Kernel_Memory}
When a \nameref{def:Process} running in \nameref{def:User} mode requests additional memory, pages are allocated from the list of free page frames maintained by the \nameref{def:Kernel}.
This list is typically populated using a \nameref{def:Page_Replacement} algorithm and most likely contains free pages scattered throughout physical memory.
Additionally, \nameref{def:Internal_Fragmentation} may result, as the process will be granted an entire page frame, even if it doesn't need all of it..

\nameref{def:Kernel} memory is often allocated from a free-memory pool \textbf{different} from the list used to satisfy ordinary user-mode processes.
There are two primary reasons for this:
\begin{enumerate}[noitemsep]
\item The \nameref{def:Kernel} requests memory for data structures of varying sizes, some of which are less than a page in size.
  As a result, the kernel must use memory conservatively and attempt to minimize waste due to \nameref{def:Fragmentation}.
  Many \nameref{def:Operating_System}s do not subject kernel code or data to the paging system.
\item Pages allocated to \nameref{def:User}-mode processes do not necessarily have to be in contiguous physical memory.
  However, certain hardware devices interact directly with physical memory, not the \nameref{def:Virtual_Memory} interface.
  Consequently, the devices may require memory residing in \textbf{physically} contiguous pages.
\end{enumerate}

We discuss 2 systems for allocating memory to \nameref{def:Kernel} \nameref{def:Process}es.
\begin{enumerate}[noitemsep]
\item \nameref{subsubsec:Buddy_System}
\item \nameref{subsubsec:Slab_Allocation}
\end{enumerate}

\subsubsection{Buddy System}\label{subsubsec:Buddy_System}
The buddy system allocates memory from a fixed-size segment consisting of \textbf{physically contiguous pages}.
Memory is allocated from this segment using a \emph{power-of-2 allocator}, which satisfies requests in units sized as a power of 2 (\SIlist{4;8;16}{\kibi{} \byte{}}, and so forth).
All requests are rounded up to the next appropriate highest power of 2.
For example, a request for \SI{11}{\kibi{} \byte{}} is satisfied with a \SI{16}{\kibi{} \byte{}} segment.

\paragraph{Benefits of the Buddy System}\label{par:Buddy_System_Benefits}
An advantage of the buddy system is how quickly adjacent buddies can be combined to form larger segments using a technique known as \emph{coalescing}.

\paragraph{Drawbacks of the Buddy System}\label{par:Buddy_System_Drawbacks}
The drawback to the buddy system is that rounding up to the next highest power of 2 will likely cause \nameref{def:Fragmentation} within allocated segments.

\subsubsection{Slab Allocation}\label{subsubsec:Slab_Allocation}
There are several components that make up the slab allocation scheme:
\begin{itemize}[noitemsep]
\item \nameref{def:Slab}
\item \nameref{def:Slab_Cache}
\item \nameref{def:Slab_Object}
\end{itemize}

\begin{definition}[Slab]\label{def:Slab}
  A \emph{slab} is made up of one or more physically contiguous pages.
\end{definition}

\begin{definition}[Cache]\label{def:Slab_Cache}
  A \emph{cache} consists of one or more \nameref{def:Slab}s.
  There is a single cache for each unique kernel data structure.

  The caches do not necessarily keep the \nameref{def:Slab_Object}s in order.
  They only guarantee that the objects will be in contiguous memory because the \nameref{def:Slab}s they map to are contiguous.

  For example, a separate cache for the data structure representing \nameref{def:Process_Descriptor}s, a separate cache for file objects, a separate cache for \nameref{def:Semaphore}s, and so forth.
\end{definition}

\begin{definition}[Object]\label{def:Slab_Object}
  \emph{Objects} that are instantiations of the kernel data structure the \nameref{def:Slab_Cache} represents.

  For example, the cache representing \nameref{def:Semaphore}s stores instances of semaphore objects, the cache representing \nameref{def:Process_Descriptor}s stores instances of process descriptor objects, and so forth.
\end{definition}

The slab-allocation algorithm uses \nameref{def:Slab_Cache}s to store \nameref{def:Kernel} objects.
When a \nameref{def:Slab_Cache} is created, a number of \nameref{def:Slab_Object}s are allocated to the \nameref{def:Slab_Cache},which are initially marked as \texttt{free}.
The number of \nameref{def:Slab_Object}s in the \nameref{def:Slab_Cache} depends on the size of the associated \nameref{def:Slab}.

For example, a \SI{12}{\kibi{} \byte{}} slab (made up of three continguous \SI{4}{\kibi{} \byte{}} pages) could store six \SI{2}{\kibi{} \byte{}} objects.
When a new object for a kernel data structure is needed, the allocator can assign any free object from the cache to satisfy the request.
The object assigned from the cache is marked as \texttt{used}.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
