\section{Laplace Transform}\label{app:Laplace_Transform}
\subsection{Laplace Transform}\label{subsec:Laplace_Transform}
\begin{definition}[Laplace Transform]\label{def:Laplace_Transform}
  The \emph{Laplace transformation} operation is denoted as $\Lapl \lbrace x(t) \rbrace$ and is defined as
  \begin{equation}\label{eq:Laplace_Transform}
    X(s) = \int\limits_{-\infty}^{\infty} x(t) e^{-st} dt
  \end{equation}
\end{definition}

\subsection{Inverse Laplace Transform}\label{subsec:Inverse_Laplace_Transform}
\begin{definition}[Inverse Laplace Transform]\label{def:Inverse_Laplace_Transform}
  The \emph{inverse Laplace transformation} operation is denoted as $\Lapl^{-1} \lbrace X(s) \rbrace$ and is defined as
  \begin{equation}\label{eq:Inverse_Laplace_Transform}
    x(t) = \frac{1}{2j \pi} \int_{\sigma-\infty}^{\sigma+\infty} X(s) e^{st} \, ds
  \end{equation}
\end{definition}

\subsection{Properties of the Laplace Transform}\label{subsec:Laplace_Transform_Properties}
\subsubsection{Linearity}\label{subsubsec:Laplace_Linearity}
The \nameref{def:Laplace_Transform} is a linear operation, meaning it obeys the laws of linearity.
This means \Cref{eq:Laplace_Linearity} must hold.
\begin{subequations}\label{eq:Laplace_Linearity}
  \begin{equation}\label{eq:Laplace_Linearity_Time}
    x(t) = \alpha_{1} x_{1}(t) + \alpha_{2} x_{2}(t)
  \end{equation}
  \begin{equation}\label{eq:Laplace_Linearity_Frequency}
    X(s) = \alpha_{1} X_{1}(s) + \alpha_{2} X_{2}(s)
  \end{equation}
\end{subequations}

\subsubsection{Time Scaling}\label{subsubsec:Laplace_Time_Scaling}
Scaling in the time domain (expanding or contracting) yields a slightly different transform.
However, this only makes sense for $\alpha > 0$ in this case.
This is seen in \Cref{eq:Laplace_Time_Scaling}.
\begin{equation}\label{eq:Laplace_Time_Scaling}
  \Lapl \bigl\lbrace x(\alpha t) \bigr\rbrace = \frac{1}{\alpha} X \left( \frac{s}{\alpha} \right)
\end{equation}

\subsubsection{Time Shift}\label{subsubsec:Laplace_Time_Shift}
Shifting in the time domain means to change the point at which we consider $t=0$.
\Cref{eq:Laplace_Time_Shifting} below holds for shifting both forward in time and backward.
\begin{equation}\label{eq:Laplace_Time_Shifting}
  \Lapl \bigl\lbrace x(t-a) \bigr\rbrace = X(s) e^{-a s}
\end{equation}

\subsubsection{Frequency Shift}\label{subsubsec:Laplace_Frequency_Shift}
Shifting in the frequency domain means to change the complex exponential in the time domain.
\begin{equation}\label{eq:Laplace_Frequency_Shift}
  \Lapl^{-1} \bigl\lbrace X(s-a) \bigr\rbrace = x(t)e^{at}
\end{equation}

\subsubsection{Integration in Time}\label{subsubsec:Laplace_Time_Integration}
Integrating in time is equivalent to scaling in the frequency domain.
\begin{equation}\label{eq:Laplace_Time_Integration}
  \Lapl \left\lbrace \int_{0}^{t} x(\lambda) \, d\lambda \right\rbrace = \frac{1}{s} X(s)
\end{equation}

\subsubsection{Frequency Multiplication}\label{subsubsec:Laplace_Frequency_Multiplication}
Multiplication of two signals in the frequency domain is equivalent to a convolution of the signals in the time domain.
\begin{equation}\label{eq:Laplace_Frequency_Multiplication}
  \Lapl \bigl\lbrace x(t) * v(t) \bigr\rbrace = X(s) V(s)
\end{equation}

\subsubsection{Relation to Fourier Transform}\label{subsubsec:Fourier_Transform_Relation}
The Fourier transform looks and behaves very similarly to the Laplace transform.
In fact, if $X(\omega)$ exists, then \Cref{eq:Fourier_Laplace_Transform_Relation} holds.
\begin{equation}\label{eq:Fourier_Laplace_Transform_Relation}
  X(s) = X(\omega) \vert_{\omega = \frac{s}{j}}
\end{equation}

\subsection{Theorems}\label{subsec:Laplace_Theorems}
There are 2 theorems that are most useful here:
\begin{enumerate}[noitemsep]
\item \nameref{thm:Laplace_Initial_Value_Theorem}
\item \nameref{thm:Laplace_Final_Value_Theorem}
\end{enumerate}

\begin{theorem}[Intial Value Theorem]\label{thm:Laplace_Initial_Value_Theorem}
  The Initial Value Theorem states that when the signal is treated at its starting time, i.e.\ $t=0^{+}$, it is the same as taking the limit of the signal in the frequency domain.
  \begin{equation*}
    x(0^{+}) = \lim\limits_{s \rightarrow \infty} s X(s)
  \end{equation*}
\end{theorem}

\begin{theorem}[Final Value Theorem]\label{thm:Laplace_Final_Value_Theorem}
  The Final Value Theorem states that when taking a signal in time to infinity, it is equivalent to taking the signal in frequency to zero.
  \begin{equation*}
    \lim\limits_{t \rightarrow \infty} x(t) = \lim\limits_{s \rightarrow 0} s X(s)
  \end{equation*}
\end{theorem}

\subsection{Laplace Transform Pairs}\label{subsec:Laplace_Transform_Pairs}
\begin{center}
  \begin{equation*}
    \renewcommand{\arraystretch}{1.4}
    \begin{array}[h!tbp]{cc}
      \toprule
      \text{Time Domain} & \text{Frequency Domain} \\
      \midrule
      x(t) & X(s) \\
      \delta(t) & 1 \\
      \delta(t-T_{0}) & e^{-sT_{0}} \\
      \UnitStep(t) & \frac{1}{s} \\
      t^{n} \UnitStep(t) & \frac{n!}{s^{n+1}} \\
      \UnitStep(t-T_{0}) & \frac{e^{-sT_{0}}}{s} \\
      e^{at} \UnitStep(t) & \frac{1}{s-a} \\
      t^{n} e^{at} \UnitStep(t) & \frac{n!}{{(s-a)}^{n+1}} \\
      \cos(bt) \UnitStep(t) & \frac{s}{s^{2} + b^{2}} \\
      \sin(bt) \UnitStep(t) & \frac{b}{s^{2} + b^{2}} \\
      e^{-at} \cos(bt) \UnitStep(t) & \frac{s+a}{{(s+a)}^{2} + b^{2}} \\
      e^{-at} \sin(bt) \UnitStep(t) & \frac{b}{{(s+a)}^{2} + b^{2}} \\
      \midrule
      r e^{-at} \cos(bt + \theta) \UnitStep(t) &
                                                 \begin{cases}
                                                   a :& \frac{sr \cos(\theta) + ar \cos(\theta) - br \sin(\theta)}{s^{2} + 2as + (a^{2} + b^{2})} \\
                                                   b :& \frac{1}{2} \left( \frac{r e^{j \theta}}{s+a-jb} + \frac{r e^{-j \theta}}{s+a+jb} \right) \\
                                                   c :& \frac{As+B}{s^{2} + 2as + c}
                                                   \begin{cases}
                                                     r &= \sqrt{\frac{A^{2}c + B^{2} - 2ABa}{c-a^{2}}} \\
                                                     \theta &= \arctan \left( \frac{Aa - B}{A \sqrt{c-a^{2}}} \right) \\
                                                   \end{cases}
                                                 \end{cases} \\
      e^{-at} \left( A \cos(\sqrt{c-a^{2}}t) + \frac{B-Aa}{\sqrt{c-a^{2}}} \sin(\sqrt{c-a^{2}}t) \right) \UnitStep(t) & \frac{As + B}{s^{2} + 2as + c} \\
      \bottomrule
    \end{array}
  \end{equation*}
\end{center}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: shared
%%% End:
