\subsection{Goals of Protection}\label{subsec:Goals_of_Protection}
System protection was originally conceived in tandem with multiprogramming \nameref{def:Operating_System}s, so that untrustworthy \nameref{def:User}s might safely:
\begin{itemize}[noitemsep]
\item Share a common logical name space
  \begin{itemize}[noitemsep]
  \item Share a directory of files
  \end{itemize}
\item Share a common physical name space
  \begin{itemize}[noitemsep]
  \item Such as memory.
  \end{itemize}
\end{itemize}

Protection can improve reliability by detecting latent errors at the interfaces between component subsystems.
Early detection of interface errors can often prevent contamination of a healthy subsystem by a malfunctioning subsystem.
Also, an unprotected resource cannot defend against use (or misuse) by an unauthorized or incompetent user.
A protection-oriented system provides means to distinguish between authorized and unauthorized usage.

The role of protection in a computer system is to provide a mechanism for the enforcement of the policies governing resource use.

Policies for resource use may vary by application, and they may change over time.
The application programmer needs to use protection mechanisms as well, to guard resources created and supported by an application subsystem against misuse.
The \nameref{def:Operating_System} should provide protection mechanisms, but application designers can use them as well in designing their own protection software.
Note that mechanisms are distinct from policies.

\textbf{Mechanisms} determine how something will be done; \textbf{policies} decide what will be done.
This separation of policy and mechanism is important for system flexibility.
This allows the same physical system to have as complex or as simple a protection scheme as required.
This separation requires general models of protection, discussed in \Cref{subsec:Access_Matrix}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
