\section{Recursion}\label{sec:Recursion}
In Haskell, there is \textbf{no} concept of iteration.
By extension, this also means that there is no \texttt{while}, \texttt{for}, and/or \texttt{do-while} looping structures.
Instead of using iteration, \nameref{def:Recursion} is used instead.

\begin{definition}[Recursion]\label{def:Recursion}
  \emph{Recursion} is the process of defining something in terms of itself, typically in a smaller amount/value.
  By ``slightly'' reducing the size of the problem through each recursive ``iteration'', the problem domain is smaller.
  Eventually, the problem domain becomes small enough that the solution is simple.
  This is called the \emph{base case}.
  Once the simple solution has been found, then all recursions can start building from that solution.

  \begin{remark}[Recursive]\label{rmk:Recursive}
    A function that uses recursion is said to be \emph{recursive}.
  \end{remark}
\end{definition}

The reason that \nameref{def:Recursion} is used instead of iteration is for multiple reasons.
\begin{itemize}[noitemsep]
\item \nameref{def:Recursion} can compute everything iteration can.
\item \nameref{def:Recursion} is inductively provable.
\item \nameref{def:Recursion} is side-effect free.
\item \nameref{def:Recursion} is a natural way to work with algebraic datatypes, lists, and the infinite structures of Functional Programming.
\end{itemize}

One of the first \nameref{rmk:Recursive} programs most people write is the factorial function.
\begin{listing}[h!tbp]
\haskellsourcefile{./EDAN40-Functional_Programming-Sections/Recursion/Code/Factorial.hs}
\caption{Factorial, Recursively Defined}
\label{lst:Recursive_Factorial}
\end{listing}

Sometimes library functions simplify the way a function is written, but many function also have simple and natural definitions using \nameref{def:Recursion}.

\begin{remark*}
  Note that throughout this section, we are regularly writing functions with the same name as functions included in the \texttt{Prelude.hs} library file.
  To prevent possible namespace collisions, functions I am writing that have the same name as \texttt{Prelude.hs} library functions will be named slightly differently, typically preceeded with \texttt{my}.
\end{remark*}

\begin{remark*}
  Throughout this section, many \nameref{rmk:Recursive} functions will be created, many of them quite inefficient.
  However, they are written this way to emphasize clarity and convey what recursion is and how to write recursive structures.
\end{remark*}

\input{./EDAN40-Functional_Programming-Sections/Recursion/List_Recursion}

\input{./EDAN40-Functional_Programming-Sections/Recursion/Multiple_Arguments}

\input{./EDAN40-Functional_Programming-Sections/Recursion/Multiple_Recursion}

\input{./EDAN40-Functional_Programming-Sections/Recursion/Mutual_Recursion}

\input{./EDAN40-Functional_Programming-Sections/Recursion/Help_Recursion}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../EDAN40-Functional_Programming-Reference_Sheet"
%%% End:
