\subsection{Directory Implementation}\label{subsec:Directory_Implementation}
The selection of directory-allocation and directory-management algorithms significantly affects the efficiency, performance, and reliability of the file system.

\subsubsection{Linear List}\label{subsubsec:Linear_List_Directory}
The simplest method of implementing a directory is to use a linear list of file names with pointers to the data blocks.
This method is simple to program but time-consuming to execute.
\begin{itemize}[noitemsep]
\item To create a new file, we must first \textbf{search} the directory to be sure that no existing file has the same name.
  Then, we add a new entry at the end of the directory.
\item To delete a file, we \textbf{search} the directory for the named file and then release the space allocated to it.
\item To reuse the directory entry, we can do one of several things:
  \begin{enumerate}[noitemsep]
  \item Mark the entry as unused (by assigning it a special name, or by including a used/unused bit).
  \item Attach it to a list of free directory entries.
  \item Copy the last entry in the directory into the freed location and to decrease the length of the directory.
  \end{enumerate}
\end{itemize}

The real disadvantage of a linear list of directory entries is that finding a file requires a linear search.
Directory information is used frequently, and users will notice if access to it is slow.

We can alleviate this problem by using a modification of a linear list.
A sorted list allows a binary search and decreases the average search time.
However, the requirement that the list be kept sorted may complicate creating and deleting files, since we may have to move substantial amounts of directory information to maintain a sorted directory.
An advantage of the sorted list is that a sorted directory listing can be produced without a separate sort step.
A more sophisticated tree data structure, such as a balanced tree, might help here.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
