\subsection{Convert I/O Requests to Hardware Operations}\label{subsec:Convert_IO_Requests_HW_Ops}
Up to this point, no explanation has been given for how the \nameref{def:Operating_System} connects an application's request to a set of wires or hardware.

In the case of a \nameref{def:File} lookup, there are 2 methods:
\begin{enumerate}[noitemsep]
\item The \nameref{def:Path_Name} also specifies the \nameref{def:Volume} or disk that the file is on.
  Windows/MS-DOS use this.
\item The device name space is incorporated into the regular \nameref{def:File_System} namespace.
  \textsc{unix} uses this.
\end{enumerate}

If the \nameref{def:Path_Name} also specifies the \nameref{def:Volume}, then the device is easy to choose from the rest of the path.
This separation makes it easy for the operating system to associate extra functionality with each device.
For instance, it is easy to invoke spooling on any files written to the printer.

If the device name space is incorporated in the regular \nameref{def:File_System} namespace, as it is in \textsc{unix}, the normal file-system name services are provided for free.
For example, if the file system provides ownership and access control to all file names, then devices also have \nameref{def:File_Owner}s and access control.
Since files are stored on devices, such an interface provides access to the I/O system at two levels.

\begin{enumerate}[noitemsep]
\item Names can be used to access the devices themselves or to access the files stored on the devices.
  Because there is no clear   separation of the device from the path the \nameref{def:Operating_System} has a mount table that associates prefixes of \nameref{def:Path_Name}s with specific device names.
  To resolve a path name, UNIX looks up the name in the mount table with the longest matching prefix; the corresponding entry in the mount table gives the device name.

\item This device name also has the form of a name in the \nameref{def:File_System} namespace.
  When \textsc{unix} looks up this name in the file-system \nameref{def:Directory} structures, it finds not an \nameref{par:Inode} number but a $\langle \mathtt{major}, \mathtt{minor} \rangle$ \nameref{def:Device_Number}.
\end{enumerate}

\begin{definition}[Device Number]\label{def:Device_Number}
  The \emph{device number} is made up of 2 parts, a \texttt{major} and \texttt{minor} number, typically represented as
  \begin{equation*}
    \langle \mathtt{major}, \mathtt{minor} \rangle
  \end{equation*}

  The \texttt{major} device number identifies a device driver that should be called to handle I/O to this device.
  The \texttt{minor} device number  is passed to the device driver to index into a device table.
  The corresponding device-table entry gives the port address or the memory-mapped address of the device controller.
\end{definition}

Modern operating systems gain significant flexibility from the multiple stages of lookup tables in the path between a request and a physical device controller.
The mechanisms that pass requests between applications and drivers are general.
Thus, we can introduce new devices and drivers into a computer without recompiling the kernel.

\subsubsection{Life Cycle of I/O Request}\label{subsubsec:IO_Request_Life_Cycle}
Here, we discuss the procedure for a \nameref{def:Blocking_Syscall} I/O request.
\begin{enumerate}[noitemsep]
\item A process issues a blocking read() system call to a file descriptor of a file that has been opened previously.
\item The system-call code in the kernel checks the parameters for correctness.
\item In the case of input, if the data are already available in the buffer cache, the data are returned to the process, and the I/O request is completed.
\item Otherwise, a physical I/O must be performed.
  The process is removed from the run queue and is placed on the wait queue for the device, and the I/O request is scheduled.
  Eventually, the I/O subsystem sends the request to the device driver.
\item The device driver allocates kernel buffer space to receive the data and schedules the I/O.
  Eventually, the driver sends commands to the device controller by writing into the device-control registers.
\item The device controller operates the device hardware to perform the data transfer.
\item The driver may poll for status and data, or it may have set the \nameref{def:Direct_Memory_Access_Controller} up to handle the transfer into kernel memory.
  Assume that the transfer is managed by a DMA controller, generating an \nameref{def:Interrupt} when the transfer completes.
\item The correct interrupt handler receives the interrupt via the \nameref{def:Interrupt_Vector} table, stores any necessary data, signals the \nameref{def:Device_Driver}, and returns from the \nameref{def:Interrupt}.
\item The device driver receives the signal, determines which I/O request has completed, determines the request’s status, and signals the kernel I/O subsystem that the request has been completed.
\item The kernel transfers data or return codes to the address space of the requesting process and moves the process from the wait queue back to the ready queue.
\item Moving the process to the ready queue unblocks the process.
  When the scheduler assigns the process to the CPU, the process resumes execution at the completion of the system call.
\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
