\subsection{Reduced Row Echelon Form}\label{subsec:RREF}
Using what we learned in \Cref{subsec:Inverse_Matrices} and \Cref{subsubsec:Inverse_Matrices_using_Elementary_Matrices}, we can solve systems of linear equations using a \nameref{def:Matrix}'s inverse.

This can be movtivated with a general example.
\begin{blackbox}
  Consider the system shown below.
  \begin{align*}
    a_{11} x_{1} + a_{12} x_{2} &= b_{1} \\
    a_{21} x_{1} + a_{22} x_{2} &= b_{2}
  \end{align*}

  We can convert this set of equations with 2 equations and 2 unknowns to a single linear equation with a single unknown.
  \textbf{Equivalently} to the system, we can write
  \begin{align*}
    \begin{pmatrix}
      a_{11} & a_{12} \\
      a_{21} & a_{22}
    \end{pmatrix}
    \begin{pmatrix}
      x_{1} \\
      x_{2}
    \end{pmatrix} &=
    \begin{pmatrix}
      a_{11} x_{1} + a_{12} x_{2} \\
      a_{21} x_{1} + a_{22} x_{2}
    \end{pmatrix} \\
    &=
      \begin{pmatrix}
        b_{1} \\
        b_{2}
      \end{pmatrix}
  \end{align*}

  The column matrices (column vectors) are technically one unknown with multiple unknown components.

  Just like with regular algebra, we can use the inverse of the items we know to solve the equation.
  \begin{align*}
    \intertext{Let}
    A &=
        \begin{pmatrix}
          a_{11} & a_{12} \\
          a_{21} & a_{22}
        \end{pmatrix} \\
    X &=
        \begin{pmatrix}
          x_{1} \\
          x_{2}
        \end{pmatrix} \\
    B &=
        \begin{pmatrix}
          b_{1} \\
          b_{2}
        \end{pmatrix} \\
    AX &= B \\
    \shortintertext{We make the assumption $A$ is invertible.}
    A^{-1} (A X) &= A^{-1} B \\
    (A^{-1} A) X &= A^{-1} B \\
    I X &= A^{-1} B
  \end{align*}

  Therefore, we can solve the system using $A^{-1}$.
\end{blackbox}

There are 3 possible cases for a solution to a system:
\begin{enumerate}[noitemsep]
\item There is no solution. This happens when one equation contradicts another.
\item There are infinitely many solutions. (\Cref{ex:System with Infinitely Many Solutions}).
\item There is exactly one solution. (\Cref{ex:Solve General System Using Matrices}).
\end{enumerate}

Now, we can illustrate with some examples.

\begin{example}[Lecture 15, Example 2]{Solve General System Using Matrices}
  Solve the system of equations shown below for all $a$, $b$, $c$?
  \begin{align*}
    x - 2y + 3z &= a \\
    -x + 0y + 2z &= b \\
    x + y - z &= c
  \end{align*}
  \tcblower{}
  First, we convert to a matrix equation.
  \begin{align*}
    \begin{pmatrix}
      1 & -2 & -3 \\
      -1 & 0 & 2 \\
      1 & 1 & -1
    \end{pmatrix}
              \begin{pmatrix}
                x \\
                y \\
                z
              \end{pmatrix} &=
                              \begin{pmatrix}
                                a \\
                                b \\
                                c
                              \end{pmatrix} \\
    \intertext{We actually already know that $A$ has an inverse. We found it in \Cref{ex:Elementary Row Operations to find Inverse}.}
    \begin{pmatrix}
      x \\
      y \\
      z \\
    \end{pmatrix} &=
                    \begin{pmatrix}
                      2 & 5 & 4 \\
                      -1 & -2 & -1 \\
                      1 & 3 & 2
                    \end{pmatrix}
                              \begin{pmatrix}
                                a \\
                                b \\
                                c
                              \end{pmatrix} \\
    &=
      \begin{pmatrix}
        2a + 5b + 4c \\
        -a - 2b - c \\
        a + 3b + 2c
      \end{pmatrix} \\
        &=
          \begin{pmatrix}
            2a \\
            -a \\
            a
          \end{pmatrix} +
    \begin{pmatrix}
      5b \\
      -2b \\
      3b
    \end{pmatrix} +
    \begin{pmatrix}
      4c \\
      -c \\
      2c
    \end{pmatrix} \\
    &= a
      \begin{pmatrix}
        2 \\
        -1 \\
        1
      \end{pmatrix}
    + b
    \begin{pmatrix}
      5 \\
      -2 \\
      3
    \end{pmatrix}
    + c
    \begin{pmatrix}
      4 \\
      -1 \\
      2
    \end{pmatrix}
  \end{align*}

  Thus, we have a single unique solution based on the values for $a$, $b$, and $c$.
\end{example}

Sometimes, we end up in a situation where we have infinitely many solutions to a system.
This is shown in \Cref{ex:System with Infinitely Many Solutions}.

\begin{example}[Lecture 15, Example 3]{System with Infinitely Many Solutions}
  Given the system below, identify a possible solution(s) for all $a$, $b$, and $c$, if there are any?
  \begin{align*}
    x + y - 3z &= a \\
    -x + y + 2z &= b \\
    z + 5y -4z &= c
  \end{align*}
  \tcblower{}
  First, convert to an \nameref{def:Augmented_Matrix} and check for an inverse.
  \begin{align*}
    A &=
        \begin{amat}{3}
          1 & 2 & -3 & a \\
          -1 & 1 & 2 & b \\
          1 & 5 & -4 & c
        \end{amat} \\
      &\grstep{\substack{r_{1}+r_{2} \\ -r_{1}+r_{3}}}
    \begin{amat}{3}
      1 & 2 & -3 & a \\
      0 & 3 & -1 & a+b \\
      0 & 3 & -1 & -a+c \\
    \end{amat}
  \end{align*}

  Rows 2 and 3 are equal, which means that the next row reduction will make one row all 0.
  This means that the coefficient matrix has no inverse.
  But, this doesn't mean that there isn't a solution, isntead, there are an infinite number of solutions.
  \begin{equation*}
    \grstep{-r_{2}+r_{3}}
    \begin{amat}{3}
      1 & 2 & -3 & a \\
      0 & 3 & -1 & a+b \\
      0 & 0 & 0 & -2a-b+c \\
    \end{amat}
  \end{equation*}

  Thus, the system will have a solution \textbf{if and only if}
  \begin{equation*}
    -2a - b + c = 0
  \end{equation*}

  If this relationship holds, then
  \begin{align*}
    x + 2y - 3z &= a \\
    3y - z &= a+b \\
  \end{align*}

  Solving this system, first for $y$ in terms of $z$, $a$, $b$, we get:
  \begin{equation*}
    y = \frac{1}{3} (a + b + z)
  \end{equation*}

  Now solving for $x$:
  \begin{align*}
    x + 2 \left( \frac{1}{3} (a + b + z) \right) - 3z &= a \\
    x &= \frac{1}{3} a + \frac{-2}{3} b + \frac{7}{3} z
  \end{align*}

  And $a$ is a free variable, so it can take any value, $z=z$.
\end{example}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../Math_333-MatrixAlg_ComplexVars-Reference_Sheet"
%%% End:
