\section{Lexical Analysis/Scanning}\label{sec:Lexical Analysis}
\begin{definition}[Lexical Analysis (Scanning)]\label{def:Lexical Analysis}
  \emph{Lexical Analysis} or \emph{Scanning} is the phase of the compilation process that reads in the source code text.
  It breaks the things it reads into \emph{tokens}.
  \begin{remark}
    \nameref{def:Lexical Analysis} \emph{ONLY} handles the reading IN of source code and the outputting of tokens.
    It \emph{DOES NOT} attach any meaning or put anything together.
  \end{remark}

  This means that these are the \textbf{\emph{ONLY}} types of errors that will be caught.
  \begin{javasource}
    int #s() {
      return 3;
    }
  \end{javasource}

  Because the \# token isn't understood by the scanner, the whole thing fails.
  The Scanner is just a simple look up device. It can only find things that it knows about.
  If it sees something that it has no clue about, it fails.
\end{definition}

There are several ways to implement a scanner.
One of the most common ways is the use of a \nameref{def:Finite State Automaton} or Finite State Machine through \nameref{def:Regular_Expression}s.

\subsection{Regular Expressions}\label{subsec:Regular_Expressions}
\begin{definition}[Regular Expression]\label{def:Regular_Expression}
  A \emph{regular expression}, sometimes called a \emph{regex} is a way to define a sequence of characters to form strings.
\end{definition}

There are 2 types of \nameref{def:Regular_Expression}s, based on the features available to make the regular expression.
\begin{enumerate}[noitemsep]
\item \nameref{def:Regex_Core_Notation}
\item \nameref{def:Regex_Extended_Notation}
\end{enumerate}

\begin{definition}[Core Notation]\label{def:Regex_Core_Notation}
  The \emph{core notation} of a \nameref{def:Regular_Expression} has a small number of features available.
  These are shown in \Cref{tab:Regex_Core_Notation}.

  \begin{table}[h!]
    \centering
    \begin{tabular}{ccc}
      \toprule
      \nameref{def:Regular_Expression} & Read As & Called \\
      \midrule
      $a$ & a & Symbol \\
      $M \vert N$ & $M$ or $N$ & Alternative \\
      $MN$ & $M$ followed by $N$ & Concatenation \\
      $\epsilon$ & The \nameref{def:Empty_String} & Epsilon \\
      $M*$ & Zero or more $M$ & Repetition (Kleene Star) \\
      $(M)$ & & Scope \\
      \bottomrule
    \end{tabular}
    \caption{\nameref{def:Regular_Expression} \nameref{def:Regex_Core_Notation}}
    \label{tab:Regex_Core_Notation}
  \end{table}
  Where $a$ is a symbol in the alphabet.
  $M$ and $N$ are \nameref{def:Regular_Expression}s.
\end{definition}

\begin{definition}[Extended Notation]\label{def:Regex_Extended_Notation}
  The \emph{extended notation} of a \nameref{def:Regular_Expression} contains all the features of the \nameref{def:Regex_Core_Notation}, and some additional features.
  These additional features \emph{can} be represented in the \nameref{def:Regex_Core_Notation}, but are confusing to read and write.
  
  The \nameref{def:Regex_Core_Notation} features are shown in \Cref{tab:Regex_Core_Notation}, the additional features added by the \nameref{def:Regex_Extended_Notation} are shown in \Cref{tab:Regex_Extended_Notation}.

  \begin{table}[h!]
    \centering
    \begin{tabular}{ccc}
      \toprule
      Extended \nameref{def:Regular_Expression} & Read As & Means \\
      \midrule
      $M+$ & One or more & $M M*$ \\
      $M$? & Optional & $\epsilon \lvert M$ \\
      $[aou]$ & One of $\ldots$ (a character class) & $a \vert o \vert u$ \\
      $[a-zA-Z]$ & & $a \vert b \vert \ldots \vert z \vert A \vert B \ldots \vert Z$ \\
      $[\wedge 0-9]$ & \multirow{2}{*}{Not} & \multirow{2}{*}{One character, but any one of those listed} \\
      Appel Notation: $\thicksim{}[0-9]$ & & \\
      ``a+b'' & The string & a \+ b \\
      \bottomrule
    \end{tabular}
    \caption{\nameref{def:Regular_Expression} \nameref{def:Regex_Extended_Notation}}
    \label{tab:Regex_Extended_Notation}
  \end{table}
\end{definition}

\subsection{Finite State Automata}\label{subsec:Finite State Automata}
Finite state automata are used for regular expressions (regex's) to determine a matching word.

\begin{definition}[Finite State Automaton]\label{def:Finite State Automaton}
  A \emph{finite state automaton} or \emph{finite state machine} is a mathematical model of computation.
  It is an abstract machine that can be in exactly one of a finite number of states at any given time.
  The FSM can change from one state to another in response to some external inputs; the change from one state to another is called a transition.
  An FSM is defined by a list of its states, its initial state, and the conditions for each transition.

  There are 2 types of finite statue automata:
  \begin{enumerate}[noitemsep, nolistsep]
  \item \nameref{def:DeterministicFiniteAutomataDFA}
  \item \nameref{def:Non-deterministicFiniteAutomataNFA}
  \end{enumerate}

  A deterministic finite state automata can be constructed to be equivalent to any non-deterministic one.

  \begin{remark}
    The plural of Finite State Automaton is Finite State Automata.
  \end{remark}
\end{definition}

\begin{definition}[Deterministic Finite Automata (DFA)]\label{def:DeterministicFiniteAutomataDFA}
  In a \emph{deterministic finite automaton} or \emph{DFA}, no two edges leaving from the same state are labeled with the same symbol.
  Additionally, there cannot be an edge that matches the empty string, $\epsilon$.
  A deterministic finite automaton will eventually terminate when it steps through all of its states necessary to reach the accepting state.
  
  The key difference between a \nameref{def:DeterministicFiniteAutomataDFA} and a \nameref{def:Non-deterministicFiniteAutomataNFA} is that you can always figure out the path that a deterministic finite automaton will take.
\end{definition}

\begin{definition}[Non-deterministic Finite Automata (NFA)]\label{def:Non-deterministicFiniteAutomataNFA}
  A \emph{non-deterministic finite automaton}, or \emph{NFA}, is one that has multiple edges leaving a single state that have the same symbol.
  It may also have special edges labeled with the empty string $\epsilon$, which is when a state is followed without ``eating'' any of the input string.
  A non-deterministic finite automaton may eventually terminate when it steps through all its states necessary to reach its accepting state.
  
  The key difference between a \nameref{def:Non-deterministicFiniteAutomataNFA} and a \nameref{def:DeterministicFiniteAutomataDFA} is that you cannot always determine the exact path that the \nameref{def:Finite State Automaton} will will take.
\end{definition}

\subsubsection{Converting a NFA to a DFA}\label{subsubsec:Convert_NFA_to_DFA}
There are a few steps for converting a non-deterministic finite automaton to a deterministic one.
\begin{enumerate}[noitemsep]
\item Start at the start state and enter it
\item Follow all the states that accept the empty string $\epsilon$ and combine them with the start state.
\item After that, read in the first character/word from the input and follow all the states that you combined.
  \begin{itemize}[noitemsep]
  \item This means that you will be following multiple states or edges at the same time.
  \end{itemize}
\item Continue doing this until you combine all the states down to a deterministic finite automaton.
  \begin{itemize}[noitemsep]
  \item You can have multiple instances of the same state, i.e., you can have state 5 in two different state bubbles, so long as the list of states inside is unique.
  \end{itemize}
\item The end states are found by taking the end states from the non-deterministic finite automaton and placing them in the deterministic finite automaton.
  \begin{itemize}[noitemsep]
  \item This means that if state 3 is and end state in the non-deterministic finite automaton, then every occurrence of state 3 in the deterministic finite automaton will be and end state.
  \end{itemize}
\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../EDAN65-Compilers-Reference_Sheet"
%%% End:
