\subsection{Basic Operators}\label{subsec:Basic_Operators}
Operators perform some operation on expressions.
This could be an arithmetic, a relational, logical, etc.\ operation.

\subsubsection{Arithmetic Operators}\label{subsubsec:Arithmetic_Operators}
These operators are for performing mathematical operations.
These are well-defined for integers and floating-point numbers.

\begin{description}[noitemsep]
\item[\cinline{+}] The addition operator.
  Works similarly for integers and floating-point numbers.

\item[\cinline{-}] The subtraction operator.
  Works similarly for integers and floating-point numbers.

\item[\cinline{*}] The multiplication operator.
  Works similarly for integers and floating-point numbers.

\item[\cinline{/}] The division operator.
  This returns the quotient of a division.
  This has a different result for integers and floating-point numbers.
  \begin{itemize}[noitemsep]
  \item Integers return the quotient of the division, but no fractional part.
  \item Floating-points return the entire remainder of the division.
  \end{itemize}

\item[\texttt{\%}] The modulo operator.
  Returns the remainder of a division operation when dividing integers.
  Note that this is only defined for integers.
\end{description}

\subsubsection{Logical}\label{subsubsec:Logical_Operators}
Logical operators work with boolean values.
\begin{description}[noitemsep]
\item[\cinline{!}] The logical NOT operator.
\item[\cinline{&&}] The logical AND operator.
  Returns \texttt{1} if and only if the left and right expressions are \textbf{BOTH} \texttt{1} at the same time.
  Otherwise, \texttt{0} is returned.
\item[\cinline{||}] The logical OR operator.
  Returns \texttt{0} if and only if both the left and right expressions are \texttt{0} at the same time.
  Otherwise, \texttt{1} is returned.
\end{description}

\subsubsection{Relational}\label{subsubsec:Relational_Operators}
These relational operators are used to define a value in relation to another.
Typically, these are used for boolean comparisons.

\begin{description}[noitemsep]
\item[\cinline{==}] The equality operator.
  Returns \texttt{1} if and only if the two expressions are equal, \texttt{0} otherwise.

\item[\cinline{!=}] The inequality operator.
  Returns \texttt{0} if and only if the two expressions are not equal, \texttt{1} otherwise.

\item[\cinline{>}] The greater-than operator.
  Returns \texttt{1} if and only if the left expression has a greater value than the right one.

\item[\cinline{>=}] The greater-than-or-equal-to operator.
  Returns \texttt{1} is and only if the left expression has a greater value or equal value than the right one.

\item[\cinline{<}] The less-than operator.
  Returns \texttt{1} if and only if the left expression has a lesser value than the right one.

\item[\cinline{<=}] The less-than-or-equal-to operator.
  Returns \texttt{1} is and only if the left expression has a lesser value or equal value than the right one.
\end{description}

\subsubsection{Assignment}\label{subsubsec:Assignment_Operators}
Unlike many other languages, in C, the assignment operator \texttt{=} is also an expression.
This means that when an assignment is performed, it also returns a value, in this case, it returns the value that was assigned to that particular name.

\begin{description}
\item[\cinline{=}] The assignment operator.
  Assigns a value to a given name.
  Returns the value of the assignment.

\item[\cinline{+=}] The add-and-assign operator.
  Takes the name on the left, adds the value on the right to the value on the left, and stores the result in the value on the left.

\item[\cinline{*=}] The multiply-and-assign operator.
  Takes the name on the left, multiplies the value on the left by to the value on the right, and stores the result in the value on the left.
\end{description}

Only \texttt{+=} and \texttt{*=} are shown, but there are similar ones defined for other \nameref{subsubsec:Arithmetic_Operators} too.

\subsubsection{Conditional Operator}\label{subsubsec:Conditional_Operator}
There is only one conditional operator, sometimes called the ternary operation or conditional expression.
It is defined like so: \cinline{bool ? true_exp : false_exp}.

\subsubsection{Bitwise Operators}\label{subsubsec:Bitwise_Operators}
Bitwise operators work on the component bits of a number.
This means they behave slightly differently than any other operator, but are not typically used in day-to-day calculations.
Usually, they are used to efficiently work with memory.

\begin{description}
\item[\cinline{&}] Bitwise AND
\item[\cinline{|}] Bitwise OR
\item[\cinline{^}] Bitwise exclusive OR (XOR)
\item[\cinline{~}] Bitwise negation, one's complement.
\item[\cinline{>>}] Bitwise SHIFT right
\item[\cinline{<<}] Bitwise SHIFT left
\end{description}

\subsection{Boolean Expressions}\label{subsec:Boolean_Expressions}
Because C is so low-level, the concept of \texttt{true} and \texttt{false} are defined as integers.

\begin{description}
\item[\cinline{0}] \texttt{false}.
\item[\cinline{1}] \texttt{true}.
  Technically, any non-zero value is considered \texttt{true}.
\end{description}

\begin{listing}[h!tbp]
\csourcefile{./C_Primer-Sections/Syntax/Code/logical-operators.c}
\caption{Logical Operators}
\label{lst:Logical_Operators}
\begin{minted}[frame=lines,linenos]{console}
$ ./a.out
1
1
0
0
1
\end{minted}
\end{listing}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../C_Primer"
%%% End:
