\section{Entropy}\label{sec:Entropy}
\input{./MMAE_320-Thermo-Sections/Entropy/Clausius_Inequality}

\subsubsection{Entropy and its Definitions}\label{subsec:Entropy_Definitions}
\begin{definition}[Entropy]\label{def:Entropy}
  \emph{Entropy} is an thermodynamically \nameref{def:Extensive_Property}.
  It is defined from the \nameref{subsec:Clausius_Inequality}, in \Cref{eq:Clausius_Inequality}.
  \begin{equation}\label{eq:Entropy}
    \Entropy = \frac{\Heat_{i}}{\Temp_{i}}
  \end{equation}

  Another equation to describe entropy is based on the concept of the number of energy states.
  \begin{equation}\label{eq:Entropy_Energy_States}
    \Entropy = \BoltzmannConstant \ln(W)
  \end{equation}
  where:
  \begin{description}[noitemsep]
  \item $\BoltzmannConstant$ is Boltzmann's Constant.
  \item $W$ is the number of energy states in the system.
    This value is usually proportional to temperature or some energy.
  \end{description}
\end{definition}

\begin{definition}[Specific Entropy]\label{def:Specific_Entropy}
  \emph{Specific Entropy}, like all the other specific properties of a state, is a thermodynamically \nameref{def:Intensive_Property}.
  Similarly to \nameref{def:Entropy}, specific entropy is defined from the \nameref{subsec:Clausius_Inequality}, in \Cref{eq:Clausius_Inequality}, but is also divided by the mass.
  \begin{equation}\label{eq:Entropy}
    \SpecificEntropy = \frac{\frac{\Heat_{i}}{\Temp_{i}}}{\Mass}
  \end{equation}
\end{definition}

There also exist \nameref{def:Cycle}s where the \nameref{def:Entropy} remains constant (which only happens if the \nameref{def:Process} is a \nameref{def:Reversible_Process}), which are called \nameref{def:Isentropic}.
These can only occur if the \nameref{def:Process} is \nameref{def:Adiabatic}.

\begin{definition}[Isentropic]\label{def:Isentropic}
  An \emph{isentropic} \nameref{def:Process}/\nameref{def:Cycle} has constant \nameref{def:Entropy}.
\end{definition}

\begin{example}[Textbook Problem 8.38]{Isentropic Ideal Gases}
  $\Mass = \SI{1}{\lbm}$ R-134a is expanded \nameref{def:Isentropic}ly in a \nameref{def:Closed_System} that goes from $\Pressure_{1} = \SI{100}{\psia}$ and $\Temp_{1} = \SI{100}{\degreeF}$ to $\Pressure_{2} = \SI{10}{\psia}$.
  Determine the total heat transfer and the work required for this \nameref{def:Process}?
  \tcblower{}
  \textbf{Concepts:} \\
  This is a \nameref{def:Closed_System}, so mass is constant, there is \textbf{no} mass flow.
  However, there can still be a heat and work flow. \\
  We are told this is an \nameref{def:Isentropic} process, so there is no change in temperature.
  \begin{align*}
    \Temp_{1} &= \Temp_{2} \\
    \Change{\Entropy} &= 0 \\
    \frac{\Change{\Heat}}{\Temp} &= 0 \\
    \Change{\Heat} &= 0
  \end{align*}

  There is a change in pressure, meaning the vessel itself expands.

  \textbf{Explore:} \\
  Pressure is \textbf{not} constant. \\
  The vessel is expanding, meaning there needs to be a change in energy.
  The energy is coming from the internal energy, $\InternalEnergy$, of the gas because the R-134a is not ``pushing'' against anything. \\
  We discovered that $\Change{\Heat} = 0$. \\
  We are not told about any work in, so assume $\Work_{In} = 0$.

  The energy balance is now:
  \begin{equation*}
    -\Work_{Out} = \Mass (\SpecificInternalEnergy_{2} - \SpecificInternalEnergy_{1}) \\
  \end{equation*}

  We can use Tables A.11E through A.13E to find values for R-134a.

  \textbf{Plan:} \\
  Using the pressures and temperatures, use the Tables to find the entropies and the internal energies.
  Remember that $\SpecificEntropy_{1} = \SpecificEntropy_{2}$, so we can find the different internal energies.

  \textbf{Solve:} \\
  Start by using Table A.13E at $\Pressure_{1} = \SI{100}{\psia}$ and $\Temp_{1} = \SI{100}{\degreeF}$ to find the initial properties of the R-134a.
  \begin{align*}
    \SpecificEntropy_{1} &= \SI{0.22902}{\btu\per\lbm\per\rankine} \\
    \SpecificInternalEnergy_{1} &= \SI{109.46}{\btu\per\lbm}
  \end{align*}

  Now, using the facts that $\SpecificEntropy_{1} = \SpecificEntropy_{2}$, and $\Pressure_{2} = \SI{10}{\psia}$, we use Table A.13E again, looking for a $\SpecificEntropy_{2}$ that is as similar to $\SpecificEntropy_{1}$ as possible.
  \begin{align*}
    \SpecificEntropy_{2} &= \SI{0.22949}{\btu\per\lbm\per\rankine} \\
    \Temp_{2} &= \SI{-29.52}{\degreeF} \\
    \SpecificInternalEnergy_{2} &= \SI{113.02}{\btu\per\lbm}
  \end{align*}

  \textbf{Validate:} \\

  \textbf{Generalize:} \\

\end{example}

\input{./MMAE_320-Thermo-Sections/Entropy/Isentropic_Efficiencies}

\subsection{Entropy Changes for Solids/Liquids}\label{subsec:Solids_Liquids_Entropy_Changes}
Gibbs Equation (\Cref{eq:Gibbs_Equation}) can be used to solve for these incompressible substances.
\begin{align*}
  \Temp d\Entropy &= d\InternalEnergy + \Pressure d\Volume \\
  d\Entropy &= \frac{d\InternalEnergy}{\Temp} + \frac{\Pressure d\Volume}{\Temp} \\
  \intertext{Because the substance is incompressible, the volume is constant.}
                  &= \frac{d\InternalEnergy}{\Temp} + \frac{\Pressure (0)}{\Temp} \\
  \int d\Entropy &= \int \frac{d\InternalEnergy}{\Temp} \\
  \intertext{Because we are under constant volume conditions, we can expand the internal energy factor out using the equation for \nameref{def:Specific_Heat} (\Cref{eq:Change_Internal_Energy_Specific_Heat_Constant_Volume}).}
                  &= \int \frac{\SpecificHeatVolume d\Temp}{\Temp} \\
  \Entropy_{2} - \Entropy_{1} &= \SpecificHeatVolume \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right)
\end{align*}

Because of the nature of incompressible substances, $\SpecificHeatVolume = \SpecificHeatPressure = \SpecificHeat$.

\begin{equation}\label{eq:Entropy_Change-Incompressible_Substances}
  \Entropy_{2} - \Entropy_{1} = \SpecificHeat \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right)
\end{equation}

\begin{example}[Textbook Problem 8.36]{Water Entropy Change}
  $\Mass = \SI{2}{\lbm}$ of water is inside a weighted piston-cylinder at $\Pressure_{1} = \SI{300}{\psia}$, which is heated.
  The volume of the cylinder is $\Volume = \SI{2.5}{\feet\cubed}$.
  The water is heated at a constant pressure until it reaches $\Temp_{2} = \SI{500}{\degreeF}$.
  Determine the water's final volume, $\Volume_{2}$ and the total change in \nameref{def:Entropy} $\Change{\Entropy}$?
  \tcblower{}
  \textbf{Concepts:} \\
  Piston-cylinders keep the \nameref{def:Process} under constant pressure. \\
  Because we are heating water, the volume of the water increases.
  How the pressure, temperature, and specific volume are related is by the \\
  Because the temperature is \textbf{not} constant, it is an \nameref{def:Irreversible_Process}.

  \textbf{Explore:} \\
  Because both the mass remains constant throughout this process, we can at least solve for the initial \nameref{def:Specific_Volume}.
  \begin{align*}
    \SpecificVolume &= \frac{\Volume}{\Mass} \\
    \SpecificVolume_{1} &= \frac{\Volume_{1}}{\Mass} \\
                    &= \frac{\SI{2.5}{\feet\cubed}}{\SI{2}{\lbm}} \\
    \SpecificVolume_{1} &= \SI{1.25}{\feet\cubed\per\lbm}
  \end{align*}

  Using the initial pressure, temperature, and specific volume, we can find the state from Table A.5E.
  This will also provide us with the entropy for the given state. \\
  We can repeat this same process for the final state too.

  \textbf{Plan:} \\
  Start with Table A.5E, and determine the state of the water. \\
  Find $\Entropy_{1}$ in the appropriate table.

  \textbf{Solve:} \\
  The initial state has $\Pressure_{1} = \SI{300}{\psia}$ and $\SpecificVolume_{1} = \SI{1.25}{\feet\cubed\per\lbm}$, so we use Table A.5E.
  From that table, we see that the water starts as a \nameref{def:Saturated_Mixture}, so we need to find the \nameref{def:Quality}.
  \begin{align*}
    \Quality_{1} &= \frac{\SpecificVolume_{1} - \Fluid{\SpecificVolume}}{\Vapor{\SpecificVolume} - \Fluid{\SpecificVolume}} \\
             &= \frac{1.25 - 0.01890}{1.5435 - 0.01890} \\
    \Quality_{1} &= 0.80749
  \end{align*}

  Now, using the initial quality, we can find the initial \nameref{def:Entropy}.
  \begin{align*}
    \SpecificEntropy_{1} &= \Fluid{\SpecificEntropy} + \Quality (\Vapor{\SpecificEntropy} - \Fluid{\SpecificEntropy}) \\
                     &= 0.58818 + 0.80749 (1.5111 - 0.58818) \\
    \SpecificEntropy_{1} &= \SI{1.33343}{\btu\per\lbm\per\rankine}
  \end{align*}

  Now, moving onto the final state.
  Because the pressure remains constant, $\Pressure_{2} = \SI{300}{\psia}$ and $\Temp_{2} = \SI{500}{\degreeF}$.
  Looking in Table A.4E we see that $\SaturatedPressure > \Pressure_{2}$ and in Table A.5E we see $\SaturatedTemp < \Temp_{2}$, which means that the steam is now a \nameref{def:Superheated_Vapor}.
  Using Table A.6E at $\Pressure_{2} = \SI{300}{\psia}$ and $\Temp_{2} = \SI{500}{\degreeF}$, we find:
  \begin{equation*}
    \SpecificEntropy_{2} = \SI{1.5706}{\btu\per\lbm\per\rankine}
  \end{equation*}

  Therefore, we can find the total change in \nameref{def:Entropy} like so:
  \begin{align*}
    \Change{\Entropy} &= \Mass \Change{\SpecificEntropy} \\
                      &= \SI{2}{\lbm} (1.5706 - 1.33343) \si{\btu\per\lbm\per\rankine} \\
    \Change{\Entropy} &= \SI{0.4744}{\btu\per\rankine}
  \end{align*}

  \textbf{Validate:} \\
  The easiest way to validate this is to plot the entropy against the pressure and look at the graph.

  \textbf{Generalize:} \\

\end{example}

\subsection{Entropy Changes for Ideal Gases}\label{subsec:Ideal_Gases_Entropy_Changes}
Gibbs Equation (\Cref{eq:Gibbs_Equation}) can again be used to solve for ideal gases.
\begin{align*}
  \Temp d\Entropy &= d\InternalEnergy + \Pressure d\Volume \\
  d\Entropy &= \frac{d\InternalEnergy}{\Temp} + \frac{\Pressure d\Volume}{\Temp} \\
  \intertext{Using the ideal gas equation \Cref{eq:Ideal_Gas_Equation} and the definition of change in internal energy for ideal gases \Cref{eq:Change_Internal_Energy_Specific_Heat_Constant_Volume}, we can ``simplify'' this.}
                  &= \frac{\SpecificHeatVolume d\Temp}{\Temp} + \frac{\GasConstant d\Volume}{\SpecificVolume} \\
  \int d\Entropy &= \int \frac{\SpecificHeatVolume d\Temp}{\Temp} + \int \frac{\GasConstant d\Volume}{\SpecificVolume} \\
  \shortintertext{Evaluating the integrals.}
  \Entropy_{2} - \Entropy_{1} &= \SpecificHeatVolume \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) + \GasConstant \ln \left( \frac{\SpecificVolume_{2}}{\SpecificVolume_{1}} \right)
\end{align*}

Thus, we can find the change in \nameref{def:Entropy} of an ideal gas if we know its \nameref{def:Specific_Volume}s using \Cref{eq:Entropy_Change-Ideal_Gases-Volume}.
\begin{equation}\label{eq:Entropy_Change-Ideal_Gases-Volume}
  \Entropy_{2} - \Entropy_{1} = \SpecificHeatVolume \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) + \GasConstant \ln \left( \frac{\SpecificVolume_{2}}{\SpecificVolume_{1}} \right)
\end{equation}

We can also expand using a slightly different formula, yielding a very similar result.
\begin{align*}
  \Temp d\Entropy &= d\Enthalpy - \SpecificVolume d\Pressure \\
  d\Entropy &= \frac{d\Enthalpy}{\Temp} - \frac{\SpecificVolume d\Pressure}{\Temp} \\
  \intertext{Using the ideal gas equation \Cref{eq:Ideal_Gas_Equation} and the definition of change in enthalpy for ideal gases \Cref{eq:Change_Internal_Energy_Specific_Heat_Constant_Pressure}, we can ``simplify'' this.}
                  &= \frac{\SpecificHeatPressure d\Temp}{\Temp} - \frac{\GasConstant d\Pressure}{\Pressure} \\
  \int d\Entropy &= \int \frac{\SpecificHeatPressure d\Temp}{\Temp} - \int \frac{\GasConstant d\Pressure}{\Pressure} \\
  \shortintertext{Evaluating the integrals.}
  \Entropy_{2} - \Entropy_{1} &= \SpecificHeatPressure \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right)
\end{align*}

Thus, we can find the change in \nameref{def:Entropy} of an ideal gas if we know its pressures using \Cref{eq:Entropy_Change-Ideal_Gases-Pressure}.
\begin{equation}\label{eq:Entropy_Change-Ideal_Gases-Pressure}
  \Entropy_{2} - \Entropy_{1} = \SpecificHeatPressure \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right)
\end{equation}

\begin{blackbox}
  \Cref{eq:Entropy_Change-Ideal_Gases-Volume,eq:Entropy_Change-Ideal_Gases-Pressure} both have the limitation that they will only work if the change in temperature is small.
  If we have to deal with large temperature differences, then we use \Cref{eq:Entropy_Change_Ideal_Gas-Large_Temp}.
\end{blackbox}

\begin{example}[Textbook Problem 8.83]{Change in Entropy of Ideal Gas}
  There is an amount of air at an initial pressure of $\Pressure_{1} = \SI{200}{\psia}$ and a temperature of $\Temp_{1} = \SI{500}{\degreeF}$.
  The air is then moved to a new vessel with a final pressure of $\Pressure_{2} = \SI{100}{\psia}$ and $\Temp_{2} = \SI{50}{\degreeF}$.
  If the \nameref{def:Specific_Heat} of the air remains constant, what is the change in \nameref{def:Entropy}, $\Change{\Entropy}$?
  \tcblower{}
  \textbf{Concepts:} \\
  The reduction in both temperature and pressure of the gas implies that the gas expands in volume. \\
  Air is an ideal gas. \\
  Constant \nameref{def:Specific_Heat}s is a little questionable, because of the large temperature change in this problem.
  Because of this, we can approximate the solution to give us an idea, or we can use the most accurate method. \\

  \textbf{Explore:} \\
  There are 2 equations we can use to find the change in \nameref{def:Entropy} of an ideal gas.
  \begin{align*}
    \Change{\Entropy} &= \SpecificHeatPressure \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right) \\
    \Change{\Entropy} &= \Entropy_{2}^{\circ} - \Entropy_{1}^{\circ} - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right)
  \end{align*}

  We can find values for $\Entropy_{1}^{\circ}$ and $\Entropy_{2}^{\circ}$ in Table A.21E.
  We also need to be careful that the pressures and temperatures are in absolute units.

  \textbf{Plan:} \\
  Solve both $\Change{\Entropy}$ equations. \\
  Use Table A.21E to solve the equation for large temperature differences.

  \textbf{Solve:} \\
  Starting with the equation we expect to be slightly wrong.
  \begin{equation*}
    \Change{\Entropy} = \SpecificHeatPressure \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right)
  \end{equation*}

  We need to find an appropriate $\SpecificHeatPressure$ value for the air. We can use Table A.2E for this.
  We also need to convert the given relative temperatures to absolute temperatures. \\
  Starting with the $\SpecificHeatPressure$ at \SI{500}{\degreeF}, we can pull the value directly from Table A.2E(b).
  \begin{equation*}
    \SpecificHeatPressure = \SI{0.248}{\btu\per\lbm\per\rankine}
  \end{equation*}

  Now for the the $\SpecificHeatPressure$ at \SI{50}{\degreeF}, we do the same.
  I chose to approximate to \SI{40}{\degreeF} for \SI{50}{\degreeF}.
  \begin{equation*}
    \SpecificHeatPressure = \SI{0.240}{\btu\per\lbm\per\rankine}
  \end{equation*}

  Now, we take the average of the two $\SpecificHeatPressure$.
  \begin{align*}
    \SpecificHeatPressure &= \frac{0.248 + 0.240}{2} \\
    \SpecificHeatPressure &= \SI{0.244}{\btu\per\lbm\per\rankine}
  \end{align*}

  Now, we can solve for the approximating equation.
  \begin{align*}
    \Change{\SpecificEntropy} &= \SpecificHeatPressure \ln \left( \frac{\Temp_{2}}{\Temp_{1}} \right) - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right) \\
                      &= 0.244 \ln \left( \frac{\SI{509.67}{\rankine}}{\SI{959.67}{\rankine}} \right) - 0.06855 \ln \left( \frac{100}{200} \right) \\
    \Change{\SpecificEntropy} &= \SI{-0.106894}{\btu\per\lbm\per\rankine}
  \end{align*}

  We can now also solve for the equation we expect to be more exact.
  We start by finding the $\SpecificEntropy^{\circ}$ values from Table A.21E.
  \begin{align*}
    \Temp_{1} &= \SI{959.67}{\rankine} \rightarrow \SpecificEntropy_{1}^{\circ} \approx \SI{0.74030}{\btu\per\lbm\per\rankine} \\
    \intertext{I approximated the temperature to \SI{960}{\rankine}.}
    \Temp_{2} &= \SI{509.67}{\rankine} \rightarrow \SpecificEntropy_{2}^{\circ} \approx \SI{0.587}{\btu\per\lbm\per\rankine} \\
    \intertext{I approximated the temperature to \SI{510}{\rankine}.}
  \end{align*}

  Using these values, we can now solve for the change in \nameref{def:Entropy}.
  \begin{align*}
    \Change{\SpecificEntropy} &= \SpecificEntropy_{2}^{\circ} - \SpecificEntropy_{1}^{\circ} - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right) \\
                      &= 0.587 - 0.74030 - 0.06855 \ln \left( \frac{100}{200} \right) \\
    \Change{\SpecificEntropy} &= \SI{-0.105785}{\btu\per\lbm\per\rankine}
  \end{align*}

  \textbf{Validate:} \\
  By solving for both $\Change{\SpecificEntropy}$, we actually ended up validating our answer as we went.
  Both numbers are quite similar to one another, so it is likely that the approximation was correct.

  The \nameref{def:Entropy} being negative is perfectly valid, because of two reasons:
  \begin{enumerate}[noitemsep]
  \item We didn't consider the entire \nameref{def:System}.
    We only considered inside the boundary of the system.
    If we included the surroundings, then we would see the net change in entropy is positive.
  \item The temperature of the gas is decreasing, reducing the overall number of possible energy states, which means the entropy decreases as well.
  \end{enumerate}

  \textbf{Generalize:} \\
  The approximation here was not significantly different than the exact calculation.
  This happened because the \nameref{def:Specific_Heat} did not significantly differ over the temperature range.
  If the temperature change were larger, then the approximation would start becoming less accurate compared to the exact equation.
\end{example}

Now, for large temperature differences, we use an equation that has already been provided for us, \Cref{eq:Entropy_Change_Ideal_Gas-Large_Temp}.
\begin{equation}\label{eq:Entropy_Change_Ideal_Gas-Large_Temp}
  \Entropy_{2} - \Entropy_{1} = \Entropy_{2}^{\circ} - \Entropy_{1}^{\circ} - \GasConstant \ln \left( \frac{\Pressure_{2}}{\Pressure_{1}} \right)
\end{equation}
\begin{description}[noitemsep]
\item $\Entropy_{2}^{\circ}$ and $\Entropy_{1}^{\circ}$ come from Table A.21 in the textbook.
\end{description}

\begin{example}[Textbook Problem 8.97]{Use Entropy Change of Ideal Gas}
  Air enters an \nameref{def:Adiabatic} nozzle at $\Pressure_{In} = \SI{60}{\psia}$, $\Temp_{In} = \SI{540}{\degreeF}$, and $\Velocity_{In} = \SI{200}{\feet\per\second}$ and exits at $\Pressure_{Out} = \SI{12}{\psia}$.
  Assuming air to be an ideal gas with variable \nameref{def:Specific_Heat}s and disregarding any irreversibilities, determine the exit velocity, $\Velocity_{Out}$ of the air?
  \tcblower{}
  \textbf{Concepts:} \\
  Because this is a nozzle, we know that $\Velocity_{Out} > \Velocity_{In}$. \\
  All the energy of the fluid is in the \nameref{def:Enthalpy} and the kinetic energy. \\
  We are told that the fluid has a variable \nameref{def:Specific_Heat}. \\
  Because we are told the the nozzle is \nameref{def:Adiabatic} and are told to ignore any irreversibilities, we can assume the nozzle is \nameref{def:Isentropic} meaning $\Entropy_{In} = \Entropy_{Out}$. \\

  \textbf{Explore:} \\
  Variable specific heats for an ideal gas under \nameref{def:Isentropic} conditions is:
  \begin{equation*}
    \Entropy_{Out}^{\circ} = \Entropy_{In}^{\circ} + \GasConstant \ln \left( \frac{\Pressure_{Out}}{\Pressure_{In}} \right)
  \end{equation*}

  We can also use the equation for relative pressures:
  \begin{equation*}
    \frac{\Pressure_{Out}}{\Pressure_{In}} = \frac{\Pressure_{Out, \Relative}}{\Pressure_{In, \Relative}}
  \end{equation*}
  where the relative pressures come from Table A.21E and are functions of temperature.

  The energy balance for a nozzle is:
  \begin{equation*}
    \FlowRate{\Mass}_{In} \left( \SpecificEnthalpy_{In} + \frac{\Velocity_{In}^{2}}{2} \right) = \FlowRate{\Mass}_{Out} \left( \SpecificEnthalpy_{Out} + \frac{\Velocity_{Out}^{2}}{2} \right)
  \end{equation*}

  \textbf{Plan:} \\
  For illustration purposes, we will use the relative pressure equation.
  We used the \nameref{def:Entropy} equation in \Cref{ex:Change in Entropy of Ideal Gas}.

  \textbf{Solve:} \\
  Using the relative pressure equation, we have:
  \begin{align*}
    \frac{\Pressure_{Out}}{\Pressure_{In}} &= \frac{\Pressure_{Out, \Relative}}{\Pressure_{In, \Relative}} \\
    \shortintertext{Rearrange the terms.}
    \Pressure_{Out, \Relative} &= \frac{\Pressure_{Out}}{\Pressure_{In}} \Pressure_{In, \Relative}
  \end{align*}

  The relative pressure is a function of temperature and can be found in Table A.21E.
  The absolute temperature of the air at the inlet is:
  \begin{align*}
    \Temp_{In} &= \SI{540}{\degreeF} + 459.67 \\
               &= \SI{999.67}{\rankine} \\
    \intertext{From Table A.21E, at \SI{1000}{\rankine}, we find a relative pressure.}
    \Pressure_{In, \Relative} &= 12.30 \\
    \SpecificEnthalpy_{In} &= \SI{240.98}{\btu\per\lbm}
  \end{align*}

  Now, we can solve for the outlet's relative pressure.
  \begin{align*}
    \Pressure_{Out, \Relative} &= \frac{\Pressure_{Out}}{\Pressure_{In}} \Pressure_{In, \Relative} \\
                               &= \left( \frac{\SI{12}{\psia}}{\SI{60}{\psia}} \right) 12.30 \\
                               &= 2.46
  \end{align*}

  Now, looking through Table A.21E, we look for an entry that is as close to the outlet's relative pressure as possible, and get the \nameref{def:Specific_Enthalpy} at that temperature. \\
  Looking at the table, the outlet air is somewhere between \SI{620}{\rankine} and \SI{640}{\rankine}, so we must \nameref{def:Interpolate}.
  \begin{align*}
    \SpecificEnthalpy_{Out} &= \SpecificEnthalpy_{620} + (2.46 - \Pressure_{\Relative, 620}) \left( \frac{\SpecificEnthalpy_{640} - \SpecificEnthalpy_{620}}{\Pressure_{\Relative, 640} - \Pressure_{\Relative, 620}} \right) \\
                          &= 148.28 + (2.46 - 2.249) \left( \frac{153.09 - 148.28}{2.514 - 2.249} \right) \\
    \SpecificEnthalpy_{Out} &= \SI{152.11}{\btu\per\lbm}
  \end{align*}

  Now, solving the energy balance equation, we can find the exit velocity of the air.
  \begin{align*}
    \FlowRate{\Mass}_{In} \left( \SpecificEnthalpy_{In} + \frac{\Velocity_{In}^{2}}{2} \right) &= \FlowRate{\Mass}_{Out} \left( \SpecificEnthalpy_{Out} + \frac{\Velocity_{Out}^{2}}{2} \right) \\
    \intertext{Nozzles are steady-flow devices ($\FlowRate{\Mass}_{In} = \FlowRate{\Mass}_{Out}$), so we can cancel out the mass flowrates.}
    \SpecificEnthalpy_{In} + \frac{\Velocity_{In}^{2}}{2} &= \SpecificEnthalpy_{Out} + \frac{\Velocity_{Out}^{2}}{2} \\
    \Velocity_{Out}^{2} &= 2\SpecificEnthalpy_{In} + \Velocity_{In}^{2} - 2\SpecificEnthalpy_{Out} \\
    \Velocity_{Out} &= \sqrt{2\SpecificEnthalpy_{In} + \Velocity_{In}^{2} - 2\SpecificEnthalpy_{Out}}
  \end{align*}

  Remember that we need to perform a unit conversion on the enthalpies.
  \begin{align*}
    \Velocity_{Out} &= \sqrt{2 (\SI{240.98}{\btu\per\lbm}) \left( \frac{\SI{25037}{\feet\squared\per\second\squared}}{\SI{1}{\btu\per\lbm}} \right) + {(\SI{200}{\feet\per\second})}^{2} - 2 (\SI{152.11}{\btu\per\lbm}) \left( \frac{\SI{25037}{\feet\squared\per\second\squared}}{\SI{1}{\btu\per\lbm}} \right)} \\
    \Velocity_{Out} &= \SI{2118.98}{\feet\per\second}
  \end{align*}
\end{example}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MMAE_320-Thermo-Reference_Sheet"
%%% End:
