\subsection{Swapping}\label{subsec:Swapping}
A \nameref{def:Process} can only be executed if it is in memory.
However, while it is \textbf{NOT} being executed, it is not required for it to be in memory.
This idea forms the basis of \nameref{def:Swapping}.

\begin{definition}[Swapping]\label{def:Swapping}
  \emph{Swapping} is the action of moving a \nameref{def:Process} out of memory to a \nameref{def:Backing_Store}, or vice versa.
\end{definition}

\begin{definition}[Backing Store]\label{def:Backing_Store}
  A \emph{backing store} is another form of storage media.
  There are typically very few requirements on the type of storage media used in a backing store, but it \textbf{MUST} be readily assessible, and be quick (not as fast as RAM, but faster than traditional file system storage).
  Typically, a fast disk is used as the backing store.

  \begin{remark}[Swap]\label{rmk:Swap}
    Sometimes the \nameref{def:Backing_Store} is called \emph{the swap}, \emph{swap-area}, \emph{swap partition}, \emph{swapfile}, etc.
    They all mean the same thing\footnote{Swap Partition and Swapfile have specify the type of the \nameref{def:Backing_Store}}.
  \end{remark}
\end{definition}

\nameref{def:Swapping} allows for the total \nameref{def:Physical_Address_Space} of \textbf{ALL} \nameref{def:Process}es to exceed the real physical memory of the system.
However, the \nameref{def:Context_Switch} time in a swapping system is fairly high.

The system maintains the ready queue consisting of all processes whose memory images are on the backing store or in memory and are ready to run.
Whenever the CPU scheduler decides to execute a process, it calls the \nameref{def:Dispatcher}.
The dispatcher checks to see where the next \nameref{def:Process} is.
\begin{itemize}[noitemsep]
\item If the process is in memory, the dispatcher will reload registers and transfer control, like normal.
\item If the process is in the swap, the dispatcher will swap the selected process into memory, reload registers, and transfer control.
\item If the process is in the swap, \textbf{and} if there is no free memory region, the dispatcher swaps out a process currently in memory and swaps in the desired process, reloads registers, and transfers control.
\end{itemize}

Clearly, it would be useful to know exactly how much memory a \nameref{def:User} \nameref{def:Process} \textbf{is} using, not simply how much it \textbf{might} be using.
This way, we only need to swap what is actually used, thereby reducing swap time.
For this method to be effective, the user \textbf{must} keep the system informed of any changes in memory requirements.
Thus, a process with dynamic memory requirements will need to issue system calls (\kernelinline{request_memory()} and \kernelinline{release_memory()}) informing the \nameref{def:Operating_System} of its changing memory needs.

\nameref{def:Swapping} is constrained by other factors as well.
\textbf{If we want to swap a process, we must be sure that it is COMPLETELY idle.}
Of particular concern is any pending I/O.
A process may be waiting for an I/O operation when we want to swap that process to the \nameref{def:Backing_Store}.
However, if the I/O is asynchronously accessing user memory for I/O buffers, then the process cannot be swapped.

There are two main solutions to this problem:
\begin{enumerate}[noitemsep]
\item Never swap a process with pending I/O.
\item Execute I/O operations only into \nameref{def:Operating_System} buffers.
  \begin{itemize}[noitemsep]
  \item Transfers between operating-system buffers and process memory only occur when the process is \textbf{swapped in}.
  \item Double buffering adds overhead.
  \item Now need to copy the data twice, once into kernel memory, then from kernel memory to user memory, before the user process can access it.
  \end{itemize}
\end{enumerate}

Standard \nameref{def:Swapping} as described above is not used in modern operating systems.
However, modified versions of swapping, are used on many systems.
One common variation, is to disable swapping until the amount of free memory (unused memory available for the operating system or processes to use) falls below a threshold amount.
Swapping is then halted when the amount of free memory increases again.
Another variation involves swapping portions of processes, rather than the entire process, to decrease swap time.
Typically, these modified forms of swapping work in conjunction with \nameref{def:Virtual_Memory}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
