\subsection{Subtype Polymorphism}\label{subsec:Subtype_Polymorphism}
\begin{definition}[Subtype Polymorphism]\label{def:Subtype_Polymorphism}
  \emph{Subtype polymorphism} can be summarized with the following phrase; ``This value has a \nameref{def:Data_Type}, and while you don't know what exact type it is, this \nameref{def:Data_Type} is a more restricted (or general) version of another \nameref{def:Data_Type} that you \textit{do} know, so therefore there are some things that you can do with it''.
\end{definition}

\begin{definition}[Subtype]\label{def:Subtype}
  A \nameref{def:Data_Type} $\SemanticType{T}$ is a \emph{subtype} of type $\SemanticType{U}$, denoted $\SemanticType{T} <: \SemanticType{U}$ or $\SemanticType{U} :> \SemanticType{T}$, if any value $v : \SemanticType{T}$ can be used in any context that requires a value of type $\SemanticType{U}$.

  Some examples of subtypes are:
  \begin{itemize}[noitemsep]
  \item $\mathtt{\SemanticType{[2 TO 5]}} <: \mathtt{\SemanticType{[2 TO 6]}}$
  \item $\mathtt{\SemanticType{[2 TO 5]}} <: \mathtt{\SemanticType{[1 TO 5]}}$
  \item $\mathtt{\SemanticType{[2 TO 5]}} <: \mathtt{\SemanticType{[1 TO 6]}}$
  \item $\mathtt{\SemanticType{[2 TO 5]}} <: \mathtt{INTEGER}$
  \end{itemize}

  This is because the supertype contains all possible values that the subtype can take, and more.
  These are all based on subset relations, meaning that typing inherits a few properties.
  \begin{propertylist}
  \item Subtyping is \emph{reflexive}. Each type \SemanticType{T} is a subtype and supertype of itself, i.e.\ $\mathtt{\SemanticType{T}} <: \mathtt{\SemanticType{T}}$.\label{prop:Subtype_Reflexive}
  \item Subtyping is \emph{transitive}. Meaning if we know $\mathtt{\SemanticType{T}} <: \mathtt{\SemanticType{U}}$ and $\mathtt{\SemanticType{U}} <: \mathtt{\SemanticType{V}}$, then $\mathtt{\SemanticType{T}} <: \mathtt{\SemanticType{V}}$ is true.\label{prop:Subtype_Transitive}
  \end{propertylist}
\end{definition}

\subsubsection{Typing Conversions}\label{subsubsec:Typing_Conversions}
\begin{definition}[Widening Conversion]\label{def:Widening_Conversion}
  Whenever we use a value of a \nameref{def:Subtype} in a place that expects a supertype, the language must perform a \emph{widening conversion}.

  \begin{remark}[Implicit]\label{rmk:Widening_Conversion_Implicit}
    Most languages make this an implicit operation, because no information is lost, making it a type-safe operation.
  \end{remark}
\end{definition}

\begin{definition}[Narrowing Conversion]\label{def:Narrowing_Conversion}
  Some languages also allow us to translate a supertype to a subtype, using a \emph{narrowing conversion}.

  \begin{remark}[Explicit]\label{rmk:Narrowing_Conversion_Explicit}
    The languages that support \nameref{def:Narrowing_Conversion}s make it an explicit command.
    This instruction may fail, or may lead to information loss, meaning it is not type-safe.
    So the language designers tend to make it an explicit operation.
  \end{remark}
\end{definition}

\subsubsection{Subtyping and Records}\label{subsubsec:Subtyping_and_Records}
Take the following 2 types, written in C, as the example for this section.
\inputminted[frame=lines,linenos]{c}{./EDAP05-Concepts_Programming_Languages-Sections/Advanced_Data_Types/Code/Subtyping_Records.c}

Any function that uses \SemanticType{R1} would also work on any record of type \SemanticType{R2}.
For these, don't think about the size of the actual record, but the possible information that can be stored and the potential operations they can have.
\begin{itemize}[noitemsep]
\item Both records can have their \texttt{x} field accessed and modified, and both of these are they same type themselves, \texttt{int}.
\item \SemanticType{R2} can have its \texttt{y} field accessed and modified.
  \begin{itemize}[noitemsep]
  \item Thus, we can do everything that \SemanticType{R1} can do with \SemanticType{R2}, making \SemanticType{R2} a \nameref{def:Subtype} of \SemanticType{R1}
  \end{itemize}
\end{itemize}

\begin{equation*}
  \mathtt{\SemanticType{R2}} <: \mathtt{\SemanticType{R1}}
\end{equation*}

\subsubsection{Subtyping and Subprograms}\label{subsubsec:Subtyping_and_Subprograms}
In a given subprogram, with \nameref{def:Formal_Parameter} inputs and some return values, there are a few ways we can use subtyping to change the operation of the subprogram.

If we need a subprogram with type $\SemanticType{T_{2}} \rightarrow \SemanticType{U_{2}}$, we can instead provide a subroutine of type $\SemanticType{T_{1}} \rightarrow \SemanticType{U_{1}}$ if $\SemanticType{T_1}$ is \textbf{less constrained} than $\SemanticType{T_{2}}$ and if $\SemanticType{U_{1}}$ is \textbf{more constrained} than $\SemanticType{U_{2}}$.
Meaning, for a subprogram, we can put \textbf{supertypes of the parameter's type} into the subprogram and it will still work.
Similarly, if we return \textbf{subtypes of the return type}, the subprogram will still work.

In other words, if we think of the function in the middle as a funnel whose opening we can widen and whose end we can narrow.

This relationship can be represented by \Cref{eq:Arrow_Rule}, called the \emph{Arrow Rule}.
\begin{equation}\label{eq:Arrow_Rule}
  \dfrac{\SemanticType{T_{1}} :> \SemanticType{T_{2}} \:\: \SemanticType{U_{1}} <: \SemanticType{U_{2}}}{\SemanticType{T_{1}} \rightarrow \SemanticType{U_{1}} <: \SemanticType{T_{2}} \rightarrow \SemanticType{U_{2}}}
\end{equation}

\begin{remark*}
  However, this \textbf{does not} hold true when the data is mutable!!
\end{remark*}

\input{./EDAP05-Concepts_Programming_Languages-Sections/Advanced_Data_Types/Type_Variance}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAP05-Concepts_Programming_Languages-Reference_Sheet"
%%% End:
