\section{Modelling}\label{sec:Modelling}
In our discussion of models, we have commonly referred to the OSI model for TCP/IP networking.
This is just one model, we have also used the queuing model to measure system performance.
But, just what is a model?
\begin{definition}[Model]\label{def:Model}
  A \emph{model} is a system that is constructed to have some certain similar properties to the system in reality.
  A model is a way to remove certain factors from the system, or limit them, so the overall system can be tested in a variety of ways.
\end{definition}

Models are used when the real-world experiments would be:
\begin{itemize}[noitemsep]
\item Too big
\item Take too much time
\item No longer exists
\item Dangerous
\item Unethical
\item Not easily reproducible
\item Difficult to analyze
\item Limiting
\end{itemize}

In this course, we will be looking at 2 major models of CSMA-based network traffic.
\begin{enumerate}[noitemsep]
\item \nameref{subsec:Kleinrock_Tobagi_Model}
\item \nameref{subsec:Bianchi_Model}
\end{enumerate}

\subsection{Model Creation}\label{subsec:Model_Creation}
\nameref{def:Model}s must be created by following the steps below.
\begin{enumerate}[noitemsep]
\item Determine the purpose of the model
\item Work out which aspects of reality need to be included, and which can be ignored or simplified
\item Choose the type of model or methodology
\item Build the model
\item Analyze the Model, Compare to reality, make predictions
\item Refine the model
\end{enumerate}

The purpose of the model \textbf{MUST} be correctly determined.
Otherwise, the whole model will not correctly model what you want it to.

\subsection{Analyzing Models}\label{subsec:Model_Analysis}
Once a model has been constructed and has been running, whatever that may mean in its context, it must be analysized and compared to reality.
\begin{enumerate}[noitemsep]
\item What is being modelled?
\item What is the purpose for developing a model?
\item What aspects of reality are needed? What simplifications and assumptions can be made?
\item What techniques are available to build the model? How can we build it?
\item How is the model evaluated?
\item Can it produce results that can be validated (compared to reality or an already-trusted model)?
\item Can we prove the validity of the model?
\item Can it make predictions that we can test?
\end{enumerate}

\subsection{Simulations}\label{subsec:Simulations}
\begin{definition}[Simulation]\label{def:Simulation}
  A \emph{simulation} is a \nameref{def:Model} that has some dynamic components, which change over time.
  These are typically implemented in software, but some systems are implemented with hardware too.

  \begin{remark}[Stochastic Simulations]\label{rmk:Stochastic_Simulations}
    Simulations do not necessarily behave the same way each time, instead they follow stochastic processes.
  \end{remark}
\end{definition}

\subsection{Kleinrock and Tobagi's Model of Packet-Switched Radio Networks}\label{subsec:Kleinrock_Tobagi_Model}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../ETSN10-Network_Architecture_Performance-Reference_Sheet"
%%% End:
