\section{Traffic Management}\label{sec:Traffic_Management}
Up to this point, we have covered ways to provide and measure network performance of various layers.
But, isolated mechanisms at different layers in the network will not achieve much on their own.
All of these mechanisms must be coordinated into a single overall strategy for \emph{traffic management}.

\subsection{Steps for Effective Traffic Management}\label{subsec:Steps_Effective_Traffic_Management}
The steps for effective traffic management are below:
\begin{enumerate}[noitemsep]
\item \textbf{Identify} applications' or flows' Quality of Service needs, and \textbf{categorize} the traffic accordingly.
\item \textbf{Enforce} traffic profiles to ensure the network has adequate resources to meet an applications' or flows' Quality of Service needs.
\item \textbf{Monitor} network performance to discover and respond to any problems that may occur.
\item \textbf{Provide} Quality of Service to meet the needs of each flow.
\end{enumerate}

\subsubsection{Identify and Categorize Traffic}\label{subsubsec:Identify_Categorize_Traffic}
Applications need a way to specify their Quality of Service needs and negotiate what kind of service they will receive from the network.
There are some specialized protocols, but they will not be covered in this course.

There are various ways to organize traffic into different classes that receive different Qualities of Service depending on the network's architecture and protocols.
\begin{itemize}[noitemsep]
\item Access Classes in 802.11 EDCA
\item \nameref{def:Bearer}s in \nameref{def:LTE}
\item \nameref{def:Network_Slicing} in \nameref{def:5G}
\end{itemize}

\subsubsection{Enforce Traffic Profiles}\label{subsubsec:Enforce_Traffic_Profiles}
For effective traffic management, the network needs a way to force traffic flows to comply with their agreed-upon traffic profiles, e.g.\ maximum data rate.
\begin{itemize}[noitemsep]
\item Admission Control: Do not allow flows into the network if there are not adequate resources available.
\item \nameref{def:Traffic_Shaping}: For example \nameref{def:Token_Bucket_Scheme}
\item \nameref{def:Flow_Control}: For example a \nameref{def:Credit_Based_Scheme}
\item Some \nameref{def:Queuing_Discipline}s ensure that a single flow cannot take more than its allowed share of bandwidth.
  \begin{itemize}[noitemsep]
  \item \nameref{def:Bit_Round_Fair_Queuing}
  \item \nameref{def:Weighted_Fair_Queuing}
  \item \nameref{def:Class_Based_Queuing}
  \end{itemize}
\end{itemize}

\subsubsection{Monitor Network Performance}\label{subsubsec:Monitor_Network_Performance}
To provide good performance, we have to be able to measure the actual performance currently achieved.
\begin{itemize}[noitemsep]
\item \nameref{def:Congestion_Control} signalling.
\item Monitoring protocols (e.g. Simple Network Management Protocol (SNMP)).
\item Signalling between routers and controller in \nameref{def:Software_Defined_Networking}.
\item Manual monitoring and response by network operators.
\end{itemize}

\subsubsection{Provide Quality of Service}\label{subsubsec:Provide_QoS}
There are many different methods for providing different types of QoS to traffic flows.
Some that are covered in this course include:
\begin{itemize}[noitemsep]
\item \nameref{def:Queuing_Discipline}s
\item Resource allocation and prioritisation in MAC protocols
\item 802.11 EDCA
\item Radio Access Network resource allocation in cellular networks
\item Reservation schemes
\item Random Early Discard (RED)
\item TCP congestion control
\end{itemize}

\subsubsection{Quality of Service Requirements}\label{subsubsec:QoS_Requirements}
QoS requirements for a given traffic flow or set of traffic flows can vary.
Some examples:
\begin{itemize}[noitemsep]
\item Real-time voice and video: low delay, high rate (depending on quality)
\item File transfer: high rate, no packet loss, no bit errors
\item Industrial control system: low delay (even hard deadlines), high reliability (packet loss and errors)
\item Sets of flows: fairness, avoiding starvation of lower priority flows
\end{itemize}

\subsection{Traffic Shaping}\label{subsec:Traffic_Shaping}
\begin{definition}[Traffic Shaping]\label{def:Traffic_Shaping}
  \emph{Traffic Shaping} is used to make \nameref{def:Stochastic_Process} more deterministic.
  During the transfer of data, we force compliance (shape the traffic) by controlling the rate at which data is sent out from the node.
  This can typically be done with \nameref{def:Token_Bucket_Scheme}s, among other options.
\end{definition}

\nameref{def:Traffic_Shaping} is typically done by determining the total capacity for all flows aggregated together.
Then, each of the flows is divided into classes that gives each flow certain properties (voice, video, data, etc.).

\begin{algorithm}[H]
  \DontPrintSemicolon{}
  \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
  \Input{Data Flows and their aggregated output capacity}
  \BlankLine{}

  \For{each flow}{
    \If{used\_capacity + (new flow) $<$ total\_capacity}{
      admit\;
    }\Else{
      no admission\;
    }
  }
\end{algorithm}

\subsubsection{Token Bucket Scheme}\label{subsubsec:Token_Bucket_Scheme}
\begin{definition}[Token Bucket Scheme]\label{def:Token_Bucket_Scheme}
  In a \emph{token bucket scheme} the transmission sender gets a set of tokens, each of which corresponds to sending a packet, byte, bit, etc.
  To send anything, you must consume one of your tokens, and these are generated over a regular interval.
\end{definition}

To find the maximum number of cells that can depart at any given time, we use \Cref{eq:Token_Bucket_Max_Cells}
\begin{equation}\label{eq:Token_Bucket_Max_Cells}
  R = \rho T + \beta
\end{equation}
where
\begin{description}[noitemsep]
\item $R$ is the Maximum number of cells that can depart from the system with a given number of tokens.
\item $\rho$ is the token generation/arrival rate.
\item $T$ is the amount of time that passes while cells are departing.
\item $\beta$ is the capacity of the token bucket.
\end{description}

To find the amount of time a transmitter can send a burst of information with a given size of token bucket, generation/arrival rate, and transmission speed, we use \Cref{eq:Token_Bucket_Max_Burst_Time}.
\begin{equation}\label{eq:Token_Bucket_Max_Burst_Time}
  T = \frac{\beta}{R_{\mathrm{Max}}-\rho}
\end{equation}
where
\begin{description}[noitemsep]
\item $T$ is the maximum amount of time a transmitter can send cells.
\item $\beta$ is the capacity of the token bucket.
\item $R_{\mathrm{Max}}$ is the maximum cell transmission rate.
\item $\rho$ is the generation/arrival of new tokens.
\end{description}

\input{./ETSN10-Network_Architecture_Performance-Sections/Traffic_Management/Congestion_Control}

\input{./ETSN10-Network_Architecture_Performance-Sections/Traffic_Management/Flow_Control}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../ETSN10-Network_Architecture_Performance-Reference_Sheet"
%%% End:
