\subsection{File Structure}\label{subsec:File_Structure}
File types can also indicate the internal structure of the file.
This point brings us to one of the disadvantages of having the operating system support multiple file structures: the resulting size of the operating system is cumbersome.
If the operating system defines five different file structures, it needs to contain the code to support these file structures.

Some operating systems impose (and support) a minimal number of file structures.
\textsc{unix} considers each file to be a sequence of bytes; no interpretation of these bits is made by the \textbf{\nameref{def:Operating_System}}.
Each application program must include its own code to interpret an input file as to the appropriate structure.
This scheme provides maximum flexibility but little support.
All operating systems must support at least one structure, the executable file, so that the system is able to load and run programs.

\subsubsection{Internal File Structure}\label{subsubsec:Internal_File_Structure}
Locating an offset within a file can be complicated for the operating system.
Disk systems typically have a well-defined block size determined by the size of a sector.
All disk I/O is performed in units of one block (physical record), and all blocks are the same size.
It is unlikely that the physical record size will exactly match the length of the desired logical record.
Because \textsc{unix} defines all files to be streams of bytes, each byte can be individually addressed by its offset from the beginning (or end) of the file.
In this case, the logical record size is 1 byte.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
