\subsection{File Types}\label{subsec:File_Types}
We need to decide whether the operating system should recognize and support \nameref{def:File} types.
If an operating system recognizes the type of a file, it can then operate on the file in reasonable ways.

A common technique for implementing \nameref{def:File} types is to include the type as part of the file name.
The name is split into two parts: the name and the extension, which are separated by a period.
This way, the user \textbf{and} the operating system can tell from the name alone what the type of a file is.

The system uses the extension to indicate the type of the \nameref{def:File} and the type of operations that can be done on that file.
Application programs also use extensions to indicate file types in which they are interested.
These extensions are not always required, so a user may specify a file without the extension, and the application will look for a file with the given name perhaps with the extension it expects.
Because these extensions are not supported by the operating system, they should be viewed as ``hints'' to the applications that operate on them.
Sometimes, the program that created the file is also stored with the \nameref{def:File_Attribute}s or the file's metadata.
In that case, the program needed to open the file does not need to be guessed.

The \textsc{unix} system uses a crude \emph{magic number} stored at the beginning of some \nameref{def:File}s to indicate roughly the type of the file.
Not all files have magic numbers, so system features cannot be based solely on this information.
\textsc{unix} also does not record the name of the creating program.
\textsc{unix} does allow file-name-extension \textbf{hints}, but they are not enforced or depended upon by the operating system.
They are meant to aid users in determining what type of contents the file contains.
Extensions can be used or ignored by a given application, but that is up to the application’s programmer.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
