\subsection{File System Mounting}\label{subsec:File_System_Mounting}
Just as a \nameref{def:File} must be opened before it is used, a \nameref{def:File_System} must be mounted before it can be available to \nameref{def:Process}es on the system.
Specifically, the directory structure may be built out of multiple \nameref{def:Volume}s, which must be mounted to make them available within the file-system name space.

\begin{definition}[Mount Point]\label{def:Mount_Point}
  A \emph{mount point} is a location within the file structure where the file system will be attached.
\end{definition}

The general steps to mount a \nameref{def:File_System} are shown below.
This scheme enables the \nameref{def:Operating_System} to traverse its directory structure, switch among file systems (possibly of various types), as appropriate.
\begin{enumerate}[noitemsep]
\item The operating system is given the name of the device and the \nameref{def:Mount_Point}.
  \begin{itemize}[noitemsep]
  \item Some operating systems require that a file system type be provided when mounting; others inspect the structures of the device and determine the type of file system.
  \item Typically, a mount point is an empty directory.
  \end{itemize}

\item The operating system verifies that the device contains a valid \nameref{def:File_System}.
  \begin{itemize}[noitemsep]
  \item It does so by asking the device driver to read the device directory
and verifying that the directory has the expected format.
\end{itemize}

\item The operating system notes in its directory structure that a file system is mounted at the specified mount point.
\end{enumerate}

Operating systems impose semantics on file system mounting to clarify functionality.
\begin{itemize}[noitemsep]
\item What if a \nameref{def:Directory} already contains \nameref{def:File}s?
  \begin{itemize}[noitemsep]
  \item A system may disallow a mount over a directory that contains files.
  \item A system may make the mounted file system available at that directory and obscure the directory’s existing files until the file system is unmounted.
    When unmounting occurs, the use of the file system and allowing access to the original files in that directory.
  \end{itemize}
\item Can a \nameref{def:File_System} be mounted multiple times at more than one \nameref{def:Mount_Point}?
  \begin{itemize}[noitemsep]
  \item A system may allow the same file system to be mounted repeatedly, at different mount points.
  \item A system may only allow one mount per file system.
  \end{itemize}
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../EDAF35-Operating_Systems-Reference_Sheet"
%%% End:
